---
title: "Freshwater Verrucocmicrobia"
author: "Edna Chiang"
date: "July 6, 2015"
output: html_document
---


### Load Libraries
```{r}
library(phyloseq)
packageVersion("phyloseq")
library(ggplot2)
library(ape)
library(vegan)
library(plyr)
library(dplyr)
library(scales)
library(grid)
library(reshape2)
library(phangorn)
library(phytools)
library(DESeq)
library(DESeq2)
library(pander)
library(ade4)
library(labdsv)
library(geiger)
#library(shiny)
#library(shinyFiles)
#library(samr)
theme_set(theme_bw())
set.seed(1)
```

### Import Data
```{r}
load("VerrucoData.RData")
```


### Visualize overall phyloseq object
```{r verruco.visualize}
# Check the sequencing depth of each sample 
sums_verruco <- data.frame(colSums(otu_table(verruco)))
colnames(sums_verruco) <- "Sample_TotalSeqs"
sums_verruco$sample <- row.names(sums_verruco)
sums_verruco <- arrange(sums_verruco, Sample_TotalSeqs)

# Create a plot of the number of sequences per sample
ggplot(sums_verruco, aes(x=reorder(sample,Sample_TotalSeqs), y = Sample_TotalSeqs)) + 
  ylab("Number of Sequences per Sample") +
  geom_bar(stat = "identity", colour="black",fill="cornflowerblue")  + xlab("Sample Name") + 
  ggtitle("Total Number of Sequences per Sample") + 
  theme(axis.text.x = element_text(colour = "black", size=6, angle=45, hjust = 1, vjust = 1))

# Look at # of reads
ggplot(data.frame(sum = sample_sums(verruco)), aes(sum)) + 
  geom_histogram(colour = "white", fill = "blue") + 
  ggtitle("Sample read counts") + 
  xlab("total sequences")

# Min, Mean, Max of sample read counts
min(sample_sums(verruco))
mean(sample_sums(verruco))
max(sample_sums(verruco))
```


### Merge Duplicates & create new phyloseq object
```{r merge.dups}
# Merges duplicates
verr.dupmerge <- merge_samples(verruco, "Duplicates")

## Fix metadata of merged duplicates 
# Grab unique lines (first entry) from verruco phyloseq object metadata
unique.idx <- !(duplicated(sample_data(verruco)$Duplicates))
duplicate.metadata <- sample_data(verruco)[unique.idx, ]

# Fix SampleID and rownames
duplicate.metadata$SampleID <- duplicate.metadata$Duplicates
rownames(duplicate.metadata) <- duplicate.metadata$Duplicates

# Assign new metadata to phyloseq object
sample_data(verr.dupmerge) <- duplicate.metadata

```

### Visualize merged dataset
```{r visualize.verr.dupmerge}
# Check the sequencing depth of each sample 
sums_dupmerge <- data.frame(rowSums(otu_table(verr.dupmerge)))
colnames(sums_dupmerge) <- "Sample_TotalSeqs"
sums_dupmerge$sample <- row.names(sums_dupmerge)
sums_dupmerge <- arrange(sums_dupmerge, Sample_TotalSeqs)

# Create a plot of the number of sequences per sample
ggplot(sums_dupmerge, aes(x=reorder(sample, Sample_TotalSeqs), y = Sample_TotalSeqs)) + 
  ylab("Number of Sequences per Sample") +
  geom_bar(stat = "identity", colour="black",fill="cornflowerblue")  + xlab("Sample Name") + 
  ggtitle("Total Number of Sequences per Sample") + 
  theme(axis.text.x = element_text(colour = "black", size=6, angle=45, hjust = 1, vjust = 1))

# Look at # of reads
ggplot(data.frame(sum = sample_sums(verr.dupmerge)), aes(sum)) + 
  geom_histogram(colour = "white", fill = "blue") +
  ggtitle("Sample read counts") + 
  xlab("total sequences")

# Mean, max and min of sample read counts
min(sample_sums(verr.dupmerge))
mean(sample_sums(verr.dupmerge))
max(sample_sums(verr.dupmerge))
```


### Trim & normalize data based on # of reads/sequence depth
```{r scale.data}
# Function to normalize data
scale_reads <- function(physeq,n){
  physeq.scale <- transform_sample_counts(physeq, function(x) {round((n*x/sum(x)))})
  #otu_table(physeq.scale) = round(otu_table(physeq.scale))
  physeq.scale = prune_taxa(taxa_sums(physeq.scale) > 0, physeq.scale)
  return(physeq.scale)
}

# Remove North Hypo samples because their reads are seriously too low (8 & 18)
dupmerge.subset <- subset_samples(physeq = verr.dupmerge, SampleID != "Z14.051.P")
dupmerge.subset <- subset_samples(physeq = dupmerge.subset, SampleID != "Z14.051.F")
#dupmerge.subset <- subset_samples(physeq = dupmerge.subset, SampleID != "Z14.016.P")
#dupmerge.subset <- subset_samples(physeq = dupmerge.subset, SampleID != "MINE.P.514")
#dupmerge.subset <- subset_samples(physeq = dupmerge.subset, SampleID != "Z14.047.P")
#dupmerge.subset <- subset_samples(physeq = dupmerge.subset, SampleID != "Sp13.ED.MM110.DD")
#dupmerge.subset <- subset_samples(physeq = dupmerge.subset, SampleID != "Z14.048.P")
#dupmerge.subset <- subset_samples(physeq = dupmerge.subset, SampleID != "Sp13.ED.MM110.SN")
#dupmerge.subset <- subset_samples(physeq = dupmerge.subset, SampleID != "Sp13.ED.MM110.SD")

### Removing or leaving in that last 10k sample doesn't make a difference in PERMANOVA/betadisper

# Normalize our dataset!
scaled <- scale_reads(dupmerge.subset,min(sample_sums(dupmerge.subset)))
  ### Scaled to 2,072 (min sample)
    # 5593 taxa
    # 226 samples
sample_data(scaled)$Lake_Type <- ordered(sample_data(scaled)$Lake_Type,
                                     levels=c("Laurentian","Estuary","Inland"))
sample_data(scaled)$Fraction <- ordered(sample_data(scaled)$Fraction,
                                     levels = c("Particle","Free","Sediment"))
sample_data(scaled)$Depth <- ordered(sample_data(scaled)$Depth,
                                  levels = c("Top", "Bottom","Sediment"))
sample_data(scaled)$Season <- ordered(sample_data(scaled)$Season,
                                   levels = c("Spring","Summer","Fall"))

# Check the sequencing depth of each sample 
sums_scaled <- data.frame(rowSums(otu_table(scaled)))
colnames(sums_scaled) <- "Sample_TotalSeqs"
sums_scaled$SampleID <- row.names(sums_scaled)
sums_scaled <- arrange(sums_scaled, Sample_TotalSeqs)

# Create a plot of the number of sequences per sample
ggplot(sums_scaled, aes(x=SampleID, y = Sample_TotalSeqs)) + 
  ylab("Number of Sequences per Sample") +
  geom_bar(stat = "identity", colour="black",fill="cornflowerblue")  + xlab("Sample Name") + 
  ggtitle("Total Number of Sequences per Sample") + 
  theme(axis.text.x = element_text(colour = "black", size=6, angle=45, hjust = 1, vjust = 1))

# Let's visualized our normalized dataset
ggplot(data.frame(sum = sample_sums(scaled)), aes(sum)) + 
  geom_histogram(colour = "black", fill = "deeppink") +
  ggtitle("Sample read counts") + 
  xlab("total sequences")
```


### Rank abundance curve of top 300 OTUs
```{r rankabund}
### Get top 500 most abundant OTUs
top300.taxa <- sort(taxa_sums(scaled), TRUE)[1:300]
top300.OTUs <- prune_taxa(names(top300.taxa), scaled)
top300.sums <- data.frame(taxa_sums(top300.OTUs))

ggplot(top300.sums,aes(x=row.names(top300.sums), y=taxa_sums.top300.OTUs.)) + 
  ylab("Number of Sequences per OTU") + scale_x_discrete(expand = c(0,0)) + 
  scale_y_continuous(expand = c(0,0)) + theme_classic() +
  geom_bar(stat="identity",colour="black",fill="darkturquoise")  + xlab("OTU Rank") + 
  ggtitle("Rank Abundance Curve of the Top 300 OTUs") + 
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank())
```



### Let's take a look at the phyla in our normalized dataset
```{r phyla.level, fig.height = 5, fig.width = 8.5}

# Merges things that have the same taxanomic rank
good_phylum <- tax_glom(scaled, taxrank = "Phylum")

# Transforms each merged phylum count to rel abun
phylum.99 = transform_sample_counts(good_phylum, function(x){(x/sum(x))*100})

# If a phylum has <1% rel abun, change that to 0%
otu_table(phylum.99)[otu_table(phylum.99) < 0.01] <- 0

# Remove anything with 0% rel abun
phylum.99 = prune_taxa(taxa_sums(phylum.99) > 0, phylum.99)

# Create rank abundance of OTUs
top5phyla.barplot <- barplot(sort(taxa_sums(phylum.99),TRUE)[1:5]/nsamples(phylum.99),
  las = 2,
  cex.axis = 0.7, 
  space = F, 
  names.arg = c("Proteobacteria", "Bacteroidetes","Actinobacteria", "Verrucomicrobia", "Cyanobacteria"), 
  col = c("khaki1", "mediumorchid3", "firebrick3", "lightslateblue", "chartreuse3"), 
  main = "Top 5 Bacterial Phyla", 
  xlab = "Phylum", 
  ylab = "Relative Abundance (%)", 
  ylim = c(0, 35), 
  las = 0
)

```

### STACKED BAR PLOT OF Lake_Type
```{r, fig.width = 7, fig.height = 8}
################# TIME TO CREATE STACKED BAR PLOT OF Lake_Type #################

# Merge samples by Lake_Type
phy.Lake_Type <- merge_samples(phylum.99, "Lake_Type")


## But this only sums them
phy.samp <- data.frame(sample_data(phy.Lake_Type))

# Average Lake_Types
phy.otu <- data.frame(otu_table(phy.Lake_Type))
phy <- data.frame(tax_table(phy.Lake_Type))
phy.otu[1,] <- (phy.otu[1,]/64)
    ## Averaging Estuary
phy.otu[2,] <- (phy.otu[2,]/123)
    ## Averaging Inland
phy.otu[3,] <- (phy.otu[3,]/32)
    ## Averaging Laurentian
#colnames(phy.otu) <- c("Proteobacteria","Actinobacteria","Cyanobacteria","Bacteroidetes","Planctomycetes","Chloroflexi","Verrucomicrobia","Armatimonadetes","NPL-UPA2","Chlorobi","Gemmatimonadetes","Acidobacteria","Candidate_division_OP3","Nitrospirae","Candidate_division_OD1","Spirochaetae","Candidate_division_OP8","Chlamydiae","Lentisphaerae","BD1-5","TM6","unclassified","Firmicutes", "Candidate_division_WS3","Fibrobacteres","Candidate_division_SR1")
tphy.otu <- t(phy.otu)
phy.otu <- otu_table(tphy.otu, taxa_are_rows = T)
phy.tax <- tax_table(phy.Lake_Type)
phy.samp <- sample_data(phy.Lake_Type)
phy.Lake_Type.merge <- merge_phyloseq(phy.tax, phy.otu, phy.samp)
    ## Makes new phyloseq object of 3 Lake_Types: Estuary, Inland, Laurentian
# Remove taxa with a relative abundane of less than 1%
otu_table(phy.Lake_Type.merge)[otu_table(phy.Lake_Type.merge) < 1] <- 0
phy.Lake_Type.merge <- prune_taxa(taxa_sums(phy.Lake_Type.merge) > 0, phy.Lake_Type.merge)



        
########### MAKE OBJECT TO BE USED IN GGPLOT ###########
phylum_long <- psmelt(phy.Lake_Type.merge)
phylum_sort <- arrange(phylum_long, Phylum)
phylum_sort$Sample <- factor(phylum_sort$Sample, levels = c("Laurentian", "Estuary", "Inland"))

# Set phylum plotting colors for 12 taxa
phylum.colors <- c(Acidobacteria="lightblue1",
                   Actinobacteria="firebrick",
                   Armatimonadetes="salmon",
                   Bacteroidetes="mediumorchid3",
                   Chlorobi="gold",
                   Chloroflexi="#ff7f00",
                   Cyanobacteria="#33a02c",
                   Firmicutes="red",
                   Nitrospirae="midnightblue",
                   Lentispherae="white",
                   Planctomycetes="deepskyblue",
                   Proteobacteria="khaki1",
                   Verrucomicrobia="lightslateblue",
                   unclassified="#662506")

# Set phylum plotting colors for 26 taxa
#phylum.colors <- c(Acidobacteria = "lightblue1", Actinobacteria = "firebrick3", Armatimonadetes = "salmon", "BD1-5" = "gold", Bacteroidetes = "mediumorchid3", "Candidate_division_OD1" = "tan", "Candidate_division_OP3" = "grey76", "Candidate_division_OP8" = "darkolivegreen1", "Candidate_division_SR1" = "cornflowerblue", "Candidate_division_WS3" = "purple4", Clamydiae = "violetred1", Chlorobi = "goldenrod", Chloroflexi = "firebrick4", Cyanobacteria = "chartreuse3", Fibrobacteres = "black", Firmicutes = "white", Gemmatimonadetes = "grey", Lentisphaerae = "yellow", "NPL-UPA2" = "hotpink", Nitrospirae = "midnightblue", Planctomycetes = "deepskyblue", Proteobacteria = "khaki1", Spirochaetae = "coral1", "TM6" = "slategray1", Verrucomicrobia = "lightslateblue", unclassified = "green4")
  
  #c(Proteobacteria = "deepskyblue", Actinobacteria = "royalblue", Cyanobacteria = "chartreuse3", Bacteroidetes = "darkorange", Planctomycetes = "mediumorchid3", Chloroflexi = "darkgreen", Verrucomicrobia = "purple4", Armatimonadetes = "red", "NPL-UPA2"="#652926", Chlorobi = "cyan2", Gemmatimonadetes = "black", Acidobacteria = "grey76", "Candidate_division_OP3" = "hotpink", Nitrospirae = "blue", "Candidate_division_OD1" = "brown", Spirochaetae = "gold3", "Candidate_division_OP8" = "goldenrod1", Chlamydiae = "violet", Lentisphaerae = "yellow1", "BD1-5" = "chartreuse", "TM6" = "olivedrab", unclassified = "grey", Firmicutes = "blue4", "Candidate_division_WS3"= "magenta", Fibrobacteres = "darkred", "Candidate_division_SR1" = "tan3" )
  
  #c(Acidobacteria = "grey26", Actinobacteria = "royalblue", Alphaproteobacteria = "plum2", Armatimonadetes = "red", Bacteroidetes = "darkorange", "BD1-5" = "chartreuse", Betaproteobacteria = "slateblue2", "Candidate_division_BRC1" = "violetred4", "Candidate_division_OD1" = "#6DDE88", "Candidate_division_OP3" = "hotpink", "Candidate_division_OP8" = "goldenrod1",  "Candidate_division_SR1" = "tan3", "Candidate_division_TM7" = "skyblue1", "Candidate_division_WS3" = "magenta", Chlamydiae="violet", Chlorobi="cyan2", Chloroflexi="darkgreen", Cyanobacteria = "chartreuse3", Deferribacteres = "slateblue3", "Deinococcus-Thermus" = "violetred", Deltaproteobacteria = "deepskyblue", Elusimicrobia = "yellow3", Epsilonproteobacteria = "lightskyblue", Fibrobacteres = "darkred", Firmicutes = "blue4", Fusobacteria = "slateblue1", Gammaproteobacteria = "steelblue4", Gemmatimonadetes="black", Lentisphaerae = "yellow1", "NPL-UPA2"="#652926", Planctomycetes = "mediumorchid3", Proteobacteria = "deepskyblue", Spirochaetae = "gold3", Tenericutes="pink", Thermotogae = "chocolate1", TM6 = "olivedrab", unclassified = "grey", Verrucomicrobia = "purple4", "WCHB1-60" = "palegreen")

# GGPLOT it

#tiff("stackbar.Lake_Type.tiff", width = 7, height = 8, units = "in", res = 500)
ggplot(phylum_sort, aes(x=Sample, y=Abundance, fill=Phylum)) +
          geom_bar(stat="identity", position="fill", color="black") +
          scale_y_continuous(labels = percent_format()) +
          scale_fill_manual(values = phylum.colors, name="Phylum") + 
          theme(axis.title.x = element_text(size=16, face = "bold"),
                axis.text.x = element_text(angle=50, colour = "black", vjust=1, hjust = 1, size=14),
                axis.text.y = element_text(colour = "black", size=14),
                axis.title.y = element_text(size=16, face="bold"),
                plot.title = element_text(size = 18, face="bold"),
                legend.title = element_text(size=14),
                legend.text = element_text(size = 13),
                legend.position="right") +
          guides(fill = guide_legend(reverse = TRUE, keywidth = 1, keyheight = 1)) + 
          xlab("Lake_Type") +
          ylab("Relative Abundance (taxa>1%)") +
          ggtitle("Lake Bacterial Community Composition")
#dev.off()

```

### LOOKING AT INTRA-VERRUCO COMMUNITY
```{r}
phy.Ver <- subset_taxa(phylum.99, Phylum == "Verrucomicrobia")
# This is the relative abundance of Verrucos in each sample
# min, max, & mean
mins <- min(sample_sums(phy.Ver))
paste(c("The minimum sample read count is", mins))
means <- mean(sample_sums(phy.Ver))
paste(c("The mean sample read count is", means))
maxs <- max(sample_sums(phy.Ver))
paste(c("The max sample read count is", maxs))

# Merge samples by Lake_Type
phy.Lake_Type.Ver <- merge_samples(phy.Ver, "Lake_Type")

# Average Lake_Types
Lake_Type.Ver <- data.frame(otu_table(phy.Lake_Type.Ver))
Lake_Type.Ver[1,] <- (Lake_Type.Ver[1,]/64)
    ## Averaging Estuary
Lake_Type.Ver[2,] <- (Lake_Type.Ver[2,]/123)
    ## Averaging Inland
Lake_Type.Ver[3,] <- (Lake_Type.Ver[3,]/32)
    ## Averaging Laurentian
#tLake_Type.Ver <- t(Lake_Type.Ver)
#tLake_Type.Ver <- data.frame(tLake_Type.Ver)
Lake_Type.Ver$Lake_Type <- rownames(Lake_Type.Ver)

# Really crappy bar plot
# Averaged relative abundance of Verrucos in each Lake_Type of lake
ggplot(data = Lake_Type.Ver, aes(x = Lake_Type, y = Otu000049)) +
  geom_bar(stat = "identity", fill = "dark grey", colour = "black") 

# Unclassifed phylum
phy.unclas <- subset_taxa(phylum.99, Phylum == "unclassified")
# Relative abundance of unclassified taxa in each sample
# min, max, & mean
mins <- min(sample_sums(phy.unclas))
paste(c("The minimum sample read count is",mins))
means <- mean(sample_sums(phy.unclas))
paste(c("The mean sample read count is",means))
maxs <- max(sample_sums(phy.unclas))
paste(c("The max sample read count is",maxs))


```


### Separate out Verrucos
```{r}
# Phyloseq object of ALL Verruco OTUs
Ver <- subset_taxa(scaled, Phylum == "Verrucomicrobia")
  # taxa = 329

# Check the sequencing depth of Verrucos each sample 
sums_Ver <- data.frame(rowSums(otu_table(Ver)))
colnames(sums_Ver) <- "Sample_TotalSeqs"
sums_Ver$sample <- row.names(sums_Ver)
sums_Ver <- arrange(sums_Ver, Sample_TotalSeqs)
summary(sums_Ver)

####  Create a plot of the number of Verruco sequences per sample
ggplot(sums_Ver, aes(x=reorder(sample, Sample_TotalSeqs), y = Sample_TotalSeqs)) + 
  ylab("Number of Sequences per Sample") +
  geom_bar(stat = "identity", colour="black",fill="cornflowerblue")  + xlab("Sample Name") + 
  ggtitle("Total Number of Sequences per Sample") + 
  theme(axis.text.x = element_text(colour = "black", size=6, angle=45, hjust = 1, vjust = 1))

```

### Intra-Verruco Rel Abund
```{r Intra.Ver}
classVer <- tax_glom(Ver, taxrank = "Class")

# Calculate relative abundance
classVer.rel = transform_sample_counts(classVer, function(x){x/sum(x)})

# If a phylum has <1% rel abun, change that to 0%
otu_table(classVer.rel)[otu_table(classVer.rel)<.01] <- 0

# Remove anything with 0% rel abun
classVer.rel = prune_taxa(taxa_sums(classVer.rel)>0,classVer.rel)

# Create rank abundance of OTUs
barplot(sort(taxa_sums(classVer.rel),TRUE)[1:20]/nsamples(classVer.rel),las=2,cex.axis=.7)



# Create dataframe to use in ggplot
classVer.df <- psmelt(classVer.rel)
classVer.df$Abundance <- classVer.df$Abundance * 100
classVer.df$Class <- ordered(classVer.df$Class, levels=c("unclassified","Verrucomicrobiae","Verrucomicrobia_Incertae_Sedis","Spartobacteria","S-BQ2-57_soil_group","Opitutae","OPB35_soil_group"))
classVer.df$Group <- ordered(classVer.df$Lake_Type, levels = c("Laurentian","Estuary","Inland"))
classVer.df$Fraction <- ordered(classVer.df$Fraction, levels = c("Particle","Free","Sediment"))
classVer.df$Depth <- ordered(classVer.df$Depth, levels = c("Top", "Bottom","Sediment"))
classVer.df$Season <- ordered(classVer.df$Season, levels = c("Spring","Summer","Fall"))


#Ver.colors <- c("OPB35_soil_group" = "gold", Opitutae = "red2", "S-BQ2-57_soil_group" = "mediumorchid3", Spartobacteria = "chartreuse3", "Verrucomicrobia_Incertae_Sedis" = "darkorange", Verrucomicrobiae = "deepskyblue", unclassified="purple4")

Ver.colors <- c("OPB35_soil_group" = "#e41a1c", Opitutae = "#377eb8", "S-BQ2-57_soil_group" = "#984ea3", Spartobacteria = "#4daf4a", "Verrucomicrobia_Incertae_Sedis" = "#ff7f00", Verrucomicrobiae = "#ffd92f", unclassified="#f1a3a9")



########### Create dataframe for unstacked bar plots (facet by Group & Fraction) ###########
classVer.rel.df <- psmelt(classVer.rel)
    ## This takes your entire phyloseq object & combines its constituents into one dataframe
sub.classVer.df <- subset(classVer.df, select = c("Sample","Class", "Abundance","Lake_Type","Fraction"))
    ## These are the factors that you'll be using to make your graph
groupfrac <- ddply(sub.classVer.df, c("Lake_Type","Fraction", "Class"), summarise,
                   n =  length(Abundance),
                   relabun = mean(Abundance),
                   sd = sd(Abundance),
                   se = sd/sqrt(n) )
    ## This takes one dataframe, applies a bunch of functions to it, and spits out the results
    ## into a new dataframe
    ## ddply ( dataframe.you.wanna.work.with, c("the factors you want to use"),
    ##         summarise (no clue what this adds to the output),
    ##         now all the functions/commands you want to apply

###### GGPLOT IT! ######
#tiff("intra.remain.lakefrac.tiff", width = 8, height = 5, units = "in", res = 500)
ggplot(groupfrac, aes(y=relabun, x = Class, fill = Class)) +
  facet_grid(Fraction ~ Lake_Type) +
  geom_bar(stat="identity", position=position_dodge()) +
  scale_fill_manual(values = Ver.colors, name = "Class") +
  theme_bw() + ggtitle("Verrucomicrobia Classes Above 0.1%") + coord_flip() +
  geom_errorbar(aes(ymin = relabun - se, ymax = relabun + se), width = 0.25) + 
  xlab("Class") + ylab("Mean Relative Abundance (%)") +
  theme(axis.title.x = element_text(face="bold", size=16),
        axis.text.x = element_text(angle=0, colour="black", size=12),
        axis.text.y = element_text(colour="black", size=12),
        axis.title.y = element_text(face="bold", size=16),
        plot.title = element_text(face="bold", size=20),
        legend.title = element_text(size=12, face="bold"),
        legend.text = element_text(size = 12),
        legend.position="none")
#dev.off()










########### Create dataframe for unstacked bar plots (facet by Season & Fraction) ###########
classVer.rel.df <- psmelt(classVer.rel)
    ## This takes your entire phyloseq object & combines its constituents into one dataframe
sub.classVer.df <- subset(classVer.df, select = c("Sample","Class", "Abundance","Season", "Fraction"))
    ## These are the factors that you'll be using to make your graph
groupfrac <- ddply(sub.classVer.df, c("Season", "Class", "Fraction"), summarise,
                   n =  length(Abundance),
                   relabun = mean(Abundance),
                   sd = sd(Abundance),
                   se = sd/sqrt(n) )
    ## This takes one dataframe, applies a bunch of functions to it, and spits out the results
    ## into a new dataframe
    ## ddply ( dataframe.you.wanna.work.with, c("the factors you want to use"),
    ##         summarise (no clue what this adds to the output),
    ##         now all the functions/commands you want to apply


###### GGPLOT IT! ######
#tiff("intra.remain.seasonfrac.tiff", width = 8, height = 5, units = "in", res = 500)
ggplot(groupfrac, aes(y=relabun, x = Class, fill = Class)) +
  facet_grid(Fraction ~ Season) +
  geom_bar(stat="identity", position=position_dodge()) +
  scale_fill_manual(values = Ver.colors, name = "Class") +
  theme_bw() + ggtitle("Verrucomicrobia Classes Above 0.1%") + coord_flip() +
  geom_errorbar(aes(ymin = relabun - se, ymax = relabun + se), width = 0.25) + 
  xlab("Class") + ylab("Mean Relative Abundance (%)") +
  theme(axis.title.x = element_text(face="bold", size=16),
        axis.text.x = element_text(angle=0, colour="black", size=12),
        axis.text.y = element_text(colour="black", size=12),
        axis.title.y = element_text(face="bold", size=16),
        plot.title = element_text(face="bold", size=20),
        legend.title = element_text(size=12, face="bold"),
        legend.text = element_text(size = 12),
        legend.position="none")
#dev.off()









########### Create dataframe for unstacked bar plots (facet by Lake_Type & Season) ###########
classVer.rel.df <- psmelt(classVer.rel)
    ## This takes your entire phyloseq object & combines its constituents into one dataframe
sub.classVer.df <- subset(classVer.df, select = c("Sample","Class", "Abundance","Season", "Lake_Type"))
    ## These are the factors that you'll be using to make your graph
groupfrac <- ddply(sub.classVer.df, c("Season", "Class", "Lake_Type"), summarise,
                   n =  length(Abundance),
                   relabun = mean(Abundance),
                   sd = sd(Abundance),
                   se = sd/sqrt(n) )
    ## This takes one dataframe, applies a bunch of functions to it, and spits out the results
    ## into a new dataframe
    ## ddply ( dataframe.you.wanna.work.with, c("the factors you want to use"),
    ##         summarise (no clue what this adds to the output),
    ##         now all the functions/commands you want to apply



###### GGPLOT IT! ######
#tiff("intra.remain.groupseason.tiff", width = 8, height = 5, units = "in", res = 500)
ggplot(groupfrac, aes(y=relabun, x = Class, fill = Class)) +
  facet_grid(Season ~ Lake_Type) +
  geom_bar(stat="identity", position=position_dodge()) +
  scale_fill_manual(values = Ver.colors, name = "Class") +
  theme_bw() + ggtitle("Verrucomicrobia Classes Above 0.1%") + coord_flip() +
  geom_errorbar(aes(ymin = relabun - se, ymax = relabun + se), width = 0.25) + 
  xlab("Class") + ylab("Mean Relative Abundance (%)") +
  theme(axis.title.x = element_text(face="bold", size=16),
        axis.text.x = element_text(angle=0, colour="black", size=12),
        axis.text.y = element_text(colour="black", size=12),
        axis.title.y = element_text(face="bold", size=16),
        plot.title = element_text(face="bold", size=20),
        legend.title = element_text(size=12, face="bold"),
        legend.text = element_text(size = 12),
        legend.position="none")
#dev.off()





########### Create dataframe for unstacked bar plots (facet by Fraction ~ Depth) ###########
classVer.rel.df <- psmelt(classVer.rel)
    ## This takes your entire phyloseq object & combines its constituents into one dataframe
sub.classVer.df <- subset(classVer.df, select = c("Sample","Class", "Abundance","Depth", "Fraction"))
    ## These are the factors that you'll be using to make your graph
groupfrac <- ddply(sub.classVer.df, c("Depth", "Class", "Fraction"), summarise,
                   n =  length(Abundance),
                   relabun = mean(Abundance),
                   sd = sd(Abundance),
                   se = sd/sqrt(n) )
    ## This takes one dataframe, applies a bunch of functions to it, and spits out the results
    ## into a new dataframe
    ## ddply ( dataframe.you.wanna.work.with, c("the factors you want to use"),
    ##         summarise (no clue what this adds to the output),
    ##         now all the functions/commands you want to apply


###### GGPLOT IT! ######
#tiff("intra.remain.depthfrac.tiff", width = 8, height = 5, units = "in", res = 500)
ggplot(groupfrac, aes(y=relabun, x = Class, fill = Class)) +
  facet_grid(Fraction ~ Depth) +
  geom_bar(stat="identity", position=position_dodge()) +
  scale_fill_manual(values = Ver.colors, name = "Class") +
  theme_bw() + ggtitle("Verrucomicrobia Classes Above 0.1%") + coord_flip() +
  geom_errorbar(aes(ymin = relabun - se, ymax = relabun + se), width = 0.25) + 
  xlab("Class") + ylab("Mean Relative Abundance (%)") +
  theme(axis.title.x = element_text(face="bold", size=16),
        axis.text.x = element_text(angle=0, colour="black", size=12),
        axis.text.y = element_text(colour="black", size=12),
        axis.title.y = element_text(face="bold", size=16),
        plot.title = element_text(face="bold", size=20),
        legend.title = element_text(size=12, face="bold"),
        legend.text = element_text(size = 12),
        legend.position="none")
#dev.off()

```



### Making phy.Ver normal
```{r}
phy.Ver.otu <- as.numeric(otu_table(phy.Ver))
hist(phy.Ver.otu,
  breaks = seq(0, 0.5, by = 0.025), 
  xlab = "Relative Abundance", 
  main = "Original Histogram")
    ## NOT NORMAL
    ## SKEWED RIGHT
qqnorm(phy.Ver.otu)
qqline(phy.Ver.otu)
    ## NOT NORMAL

# Test for normality
shapiro.test(phy.Ver.otu)
    ## DEF not normal (p-value = 0)
ks.test(phy.Ver.otu, "pnorm", mean = mean(phy.Ver.otu), sd = sd(phy.Ver.otu))
    # NOT NORMAL (p-value = 0.01615)

##### Log Transformation #####
log.ver <- log(phy.Ver.otu)
hist(log.ver, breaks=18, main = "Log-Transformed Histogram", xlab = "Log(Relative Abundance)")
    ## LOOKS BEAUTIFUL!
qqnorm(log.ver)
qqline(log.ver)
    ## Looks slightly better...
shapiro.test(log.ver)
    ## p-value = 0.08386
    ## Barely non-significant...I want a bigger p-value
ks.test(log.ver, "pnorm", mean=mean(log.ver), sd=sd(log.ver))
    ## p-value = 0.5135
    ## This is A LOT better
############ The histogram looks great, & both shapiro & ks come out not significant
############ Although shapiro is close-ish to the edge, I think it's okay based on the histogram




####### Repeat for phy.Ver.water #######
phy.Ver.water.otu <- as.numeric(otu_table(phy.Ver.water))
hist(phy.Ver.water.otu, 
  breaks = seq(0, .5, by = 0.025), 
  xlab = "Relative Abundance", 
  main = "Original Histogram"
)
    ## NOT NORMAL
    ## SKEWED RIGHT
qqnorm(phy.Ver.water.otu)
qqline(phy.Ver.water.otu)
    ## NOT NORMAL

# Test for normality
shapiro.test(phy.Ver.water.otu)
    ## DEF not normal (p-value = 0)
ks.test(phy.Ver.water.otu, "pnorm", mean = mean(phy.Ver.water.otu), sd = sd(phy.Ver.water.otu))
    # NOT NORMAL (p-value = 0.0275)

##### Log Transformation #####
log.ver.water <- log(phy.Ver.water.otu)
hist(log.ver.water, breaks = 18, main = "Log-Transformed Histogram", xlab = "Log(Relative Abundance)")
    ## LOOKS BEAUTIFUL!
qqnorm(log.ver.water)
qqline(log.ver.water)
    ## Looks slightly better...
shapiro.test(log.ver.water)
    ## p-value = 0.05037
    ## Barely non-significant...I want a bigger p-value
ks.test(log.ver.water, "pnorm", mean = mean(log.ver.water), sd = sd(log.ver.water))
    ## p-value = 0.5383
    ## This is A LOT better




################ Import normal data in phy.Ver phyloseq object ################
phy.Ver.otu <- data.frame(otu_table(phy.Ver))
phy.Ver.otu[1, ] <- log.ver

phy.Ver.water.otu <- data.frame(otu_table(phy.Ver.water))
phy.Ver.water.otu[1, ] <- log.ver.water

################# THIS IS THE VERRUCO PHYLUM PHYLOSEQ OBJECT YOU'LL USE IN THE FUTURE #################
otu <- otu_table(phy.Ver.otu, taxa_are_rows = TRUE)
tax <- tax_table(phy.Ver)
samp <- sample_data(phy.Ver)
normal.Ver <- merge_phyloseq(tax, otu, samp)


```


### Box-And-Whisker Plots
```{r}
######## YOU'RE USING THE PHY.VER.OTU DATAFRAME WITH THE NORMAL DATA ########
tphy.Ver.otu <- t(phy.Ver.otu)
phy.Ver.samp.norm <- data.frame(sample_data(phy.Ver))
phy.Ver.samp.norm$Relative_Abundance <- tphy.Ver.otu[,1]
phy.Ver.samp.norm$Lake_Type <- ordered(phy.Ver.samp.norm$Lake_Type, levels=c("Laurentian","Estuary","Inland"))
phy.Ver.samp.norm$Lake <- ordered(phy.Ver.samp.norm$Lake, levels=c("Michigan", "Muskegon", "BigSeven", "Bishop", "Bruin", "Cedar", "Halfmoon", "Heron", "Independence", "Lobdell", "North", "Sand", "Whitmore", "Woodland"))
phy.Ver.samp.norm$Fraction <- ordered(phy.Ver.samp.norm$Fraction, levels = c("Particle", "Free", "Sediment"))
phy.Ver.samp.norm$Depth <- ordered(phy.Ver.samp.norm$Depth, levels = c("Top", "Bottom", "Sediment"))

phy.Ver.otu.box <- data.frame(otu_table(phy.Ver))
tphy.Ver.box <- t(phy.Ver.otu.box)
phy.Ver.samp <- data.frame(sample_data(phy.Ver))
phy.Ver.samp$Relative_Abundance <- tphy.Ver.box[,1]
phy.Ver.samp$Lake_Type <- ordered(phy.Ver.samp$Lake_Type, levels=c("Laurentian","Estuary","Inland"))
phy.Ver.samp$Lake <- ordered(phy.Ver.samp$Lake, levels=c("Michigan", "Muskegon", "BigSeven", "Bishop", "Bruin", "Cedar", "Halfmoon", "Heron", "Independence", "Lobdell", "North", "Sand", "Whitmore", "Woodland"))
phy.Ver.samp$Fraction <- ordered(phy.Ver.samp$Fraction, levels = c("Particle", "Free", "Sediment"))
phy.Ver.samp$Depth <- ordered(phy.Ver.samp$Depth, levels = c("Top", "Bottom", "Sediment"))



################### BOXPLOT BY Lake_Type & FRACTION ###################
tiff("boxplot.tiff", width = 7, height = 5, units = "in", res = 500)
boxplot(phy.Ver.samp$Relative_Abundance ~ phy.Ver.samp$Lake_Type + phy.Ver.samp$Fraction, xlab = "Sample", ylab = "Relative Abundance", main = "Verrucomicrobia Abundance", names = rep("",9))
dev.off()

Lake_Type.box <- boxplot(phy.Ver.samp$Relative_Abundance ~ phy.Ver.samp$Lake_Type, xlab = "Lake_Type", ylab = "Relative Abundance")
pairwise.t.test(x = phy.Ver.samp.norm$Relative_Abundance, g=phy.Ver.samp.norm$Lake_Type, p.adjust="bonferroni")
    ## SIGNIFICANT btwn Estuary/Laurentian & Inland/Laurentian
    ## NOT SIGNIFICANT btwn Estuary & Inland

lake.box <- boxplot(phy.Ver.samp$Relative_Abundance ~ phy.Ver.samp$Lake, data = phy.Ver.samp, xlab = "Lake Type", ylab = "Relative Abundance", main = "Lake Type", ylim = c(0,0.45))
anova(lm(phy.Ver.samp$Relative_Abundance ~ phy.Ver.samp$Lake))
    ## Significant
pairwise.t.test(x = phy.Ver.samp.norm$Relative_Abundance, g=phy.Ver.samp.norm$Lake, p.adjust="bonferroni")
    ## SIGNIFICANT btwn Michigan/Muskegon, Michigan/Bruin, Michigan/Heron, Michigan/Independence, Michigan/North, Michigan/Sand, Michigan/Whitmore;; Michigan/Lobdell is on the cusp
    ## SIGNIFICANT btwn Muskegon/Michigan, Muskegon/North
    ## SIGNIFICANT btwn North & Michigan, Cedar

fraction.box <- boxplot(phy.Ver.samp$Relative_Abundance ~ phy.Ver.samp$Fraction, xlab = "Fraction", ylab = "Relative Abundance", main = "Fraction")
pairwise.t.test(x = phy.Ver.samp.norm$Relative_Abundance, g=phy.Ver.samp.norm$Fraction, p.adjust="bonferroni")
    ## SIGNIFICANT between Sed/Part & Sed/Free
    ## NOT SIGNIFICANT btwn Part & Free

depth.box <- boxplot(phy.Ver.samp$Relative_Abundance ~ phy.Ver.samp$Depth, xlab = "Depth", ylab = "Relative Abundance", main = "Depth")
pairwise.t.test(x = phy.Ver.samp.norm$Relative_Abundance, g=phy.Ver.samp.norm$Depth, p.adjust="bonferroni")
    ## SIGNIFICANT between Sed/Top & Sed/Bottom
    ## NOT SIGNIFICANT btwn Top & Bottom

date.box <- boxplot(phy.Ver.samp$Relative_Abundance ~ phy.Ver.samp$Season, xlab = "Season", ylab = "Relative Abundance", main = "Season")
pairwise.t.test(x = phy.Ver.samp.norm$Relative_Abundance, g=phy.Ver.samp.norm$Season, p.adjust="bonferroni")
    ## NOT SIGNIFICANT (p-value=1)

origin.box <- boxplot(phy.Ver.samp$Relative_Abundance ~ phy.Ver.samp$Origin, xlab = "Season", ylab = "Relative Abundance", main = "Season")
pairwise.t.test(x = phy.Ver.samp.norm$Relative_Abundance, g=phy.Ver.samp.norm$Origin, p.adjust="bonferroni")
    ## SIGNIFICANT (p-value = 0.001)


############################## Just water column samples (NO SEDIMENT) ##############################
######## YOU'RE USING THE PHY.VER.WATER.OTU DATAFRAME WITH THE NORMAL DATA ########
phy.Ver.water.otu <- data.frame(otu_table(phy.Ver.water))
tphy.Ver.water.otu <- t(phy.Ver.water.otu)
phy.Ver.water.samp <- data.frame(sample_data(phy.Ver.water))
phy.Ver.water.samp$Relative_Abundance <- tphy.Ver.water.otu[,1]
phy.Ver.water.samp$Lake_Type <- ordered(phy.Ver.water.samp$Lake_Type, levels=c("Laurentian","Estuary","Inland"))
phy.Ver.water.samp$Lake <- ordered(phy.Ver.water.samp$Lake, levels=c("Michigan", "Muskegon", "BigSeven", "Bishop", "Bruin", "Cedar", "Halfmoon", "Heron", "Independence", "Lobdell", "North", "Sand", "Whitmore", "Woodland"))
phy.Ver.water.samp$Fraction <- ordered(phy.Ver.water.samp$Fraction, levels = c("Particle", "Free"))
phy.Ver.water.samp$Depth <- ordered(phy.Ver.water.samp$Depth, levels = c("Top", "Bottom"))

### By Lake_Type ###
boxplot(phy.Ver.water.samp$Relative_Abundance ~ phy.Ver.water.samp$Lake_Type, xlab = "Lake Type", ylab = "Relative Abundance", main = "Lake Type")
pairwise.t.test(x = phy.Ver.water.samp$Relative_Abundance, g=phy.Ver.water.samp$Lake_Type, p.adjust="bonferroni")
    ## SIGNIFICANT btwn Estuary/Laurentian & Inland/Laurentian
    ## NOT SIGNIFICANT btwn Estuary & Inland

### By Lake ###
boxplot(phy.Ver.water.samp$Relative_Abundance ~ phy.Ver.water.samp$Lake, data = phy.Ver.samp, xlab = "Lake Type", ylab = "Relative Abundance", main = "Lake Type", ylim = c(0,0.45))
anova(lm(phy.Ver.water.samp$Relative_Abundance ~ phy.Ver.water.samp$Lake))
    ## Significant
pairwise.t.test(x = phy.Ver.water.samp$Relative_Abundance, g=phy.Ver.water.samp$Lake, p.adjust="bonferroni")
    ## SIGNIFICANT btwn Michigan/Muskegon, Michigan/Bruin, Michigan/Heron, Michigan/Independence, Michigan/North, Michigan/Sand, Michigan/Whitmore;; Michigan/Lobdell is on the cusp
    ## SIGNIFICANT btwn Muskegon/Michigan
    ## SIGNIFICANT btwn North & Michigan, BigSeven, Bishop, Bruin, Cedar

### By Fraction ###
boxplot(phy.Ver.water.samp$Relative_Abundance ~ phy.Ver.water.samp$Fraction, xlab = "Fraction", ylab = "Relative Abundance", main = "Fraction")
pairwise.t.test(x = phy.Ver.water.samp$Relative_Abundance, g=phy.Ver.water.samp$Fraction, p.adjust="bonferroni")
    ## NOT SIGNIFICANT btwn Part & Free

### By Depth ###
boxplot(phy.Ver.water.samp$Relative_Abundance ~ phy.Ver.water.samp$Depth, xlab = "Depth", ylab = "Relative Abundance", main = "Depth")
pairwise.t.test(x = phy.Ver.water.samp$Relative_Abundance, g=phy.Ver.water.samp$Depth, p.adjust="bonferroni")
    ## NOT SIGNIFICANT btwn Top & Bottom


################### REMOVING SED DOESN'T HAVE AN EFFECT ON SIG BY Lake_Type OR LAKE ###################
```

### Verruco Relative Abundaunce to test normality
```{r}
# Transforms each merged phylum count to rel abun
Ver.rel <- transform_sample_counts(Ver, function(x){x/sum(x)})

# If a phylum has <1% rel abun, change that to 0%
otu_table(Ver.rel)[otu_table(Ver.rel)<.01] <- 0

# Remove anything with 0% rel abun
Ver.rel = prune_taxa(taxa_sums(Ver.rel)>0,Ver.rel)
  # Remove taxa = 135
  # Remain taxa = 208

# Create rank abundance of OTUs
barplot(sort(taxa_sums(Ver.rel),TRUE)[1:20]/nsamples(Ver.rel),las=2,cex.axis=.7)
    ## 1 = OPB35_soil_group
    ## 2 = OPB35_soil_group
    ## 3 = Opitutae - vadinHA64
    ## 4 = Spartobacteria - Chthoniobacterales - FukuN18_freshwater_Lake_Type
    ## 5 = Verrucomicrobiae - Verrucomicrobiales - Verrucomicrobiaceae

Ver.otu <- as.numeric(otu_table(Ver.rel))
hist(Ver.otu)
    ## So skewed that I can't even get a legit histogram from it
qqnorm(Ver.otu)
qqline(Ver.otu)
    ## Looks exponentially distributed

# Test for normality
shapiro.test(Ver.otu)
    ## Too many samples
ks.test(Ver.otu, "pnorm", mean=mean(Ver.otu), sd=sd(Ver.otu))
    # p-value = 0
    # Not normal
ks.test(Ver.otu, "pexp")
    ## p-value = 9
    ## not exponentially distributed
Ver.otu <- Ver.otu + 1

#### Log Transformation ####
Ver.log <- log(Ver.otu)
hist(Ver.log)
    ## Still not normal
    ## Skewed to the right
qqnorm(Ver.log)
qqline(Ver.log)
ks.test(Ver.log, "pnorm", mean=mean(Ver.otu), sd=sd(Ver.otu))
    ## p=value = 0
    ## Not normal

#################### Verruco Intra-Phylum Rel Abun is NOT NORMAL ####################

```



### NMDS (Bray-Curtis)
```{r}
nmds.bray <- ordinate(
  physeq = Ver.rel, 
  method = "NMDS", 
  distance = "bray"
)
    # Stress = 0.2112829

### Color = Depth, Shape = Lake_Type
plot_ordination(
  physeq = Ver.rel,
  ordination = nmds.bray,
  color = "Depth",
  shape = "Lake_Type") + 
  scale_color_manual(values = c("indianred1","green3","mediumpurple2", "skyblue")) +
  scale_shape_manual(values = c(15,17,16)) +
  theme(axis.text.x = element_blank(), 
        axis.text.y = element_blank(),
        axis.title.x = element_text(face="bold", size=14),
        axis.title.y = element_text(face="bold", size=14),
        legend.title = element_text(size=12, face="bold"),
        legend.text = element_text(size=12),
        axis.ticks = element_blank(),
        plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm"), #top, right, bottom, left
        plot.title = element_text(size = 16, face="bold")) +
  guides(color = guide_legend(order=1, override.aes=list(size=3)),
         shape = guide_legend(order=2, override.aes=list(size=3))) 



### Color = Lake_Type, Shape = Depth
bray.nmds.Lake_Type.depth <- plot_ordination(
  physeq = Ver.rel,
  ordination = nmds.bray,
  color = "Lake_Type",
  shape = "Depth") + 
  scale_color_manual(values = c("indianred1","deepskyblue","mediumpurple2")) +
  ggtitle("Bray-Curtis Dissimilarity") +
  theme(axis.text.x = element_blank(), 
        axis.text.y = element_blank(),
        axis.title.x = element_text(face="bold", size=14),
        axis.title.y = element_text(face="bold", size=14),
        legend.title = element_text(size=12, face="bold"),
        legend.text = element_text(size=12),
        axis.ticks = element_blank(),
        plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm"), #top, right, bottom, left
        plot.title = element_text(size = 16, face="bold")) +
  guides(color = guide_legend(order=1, override.aes=list(size=3)),
         shape = guide_legend(order=2, override.aes=list(size=3))) 



### Color = Fraction, Shape = Lake_Type
plot_ordination(
  physeq = Ver.rel,
  ordination = nmds.bray,
  color = "Fraction",
  shape = "Depth") + 
  scale_color_manual(values = c("indianred1","deepskyblue","mediumpurple2")) +
  theme(axis.text.x = element_blank(), 
        axis.text.y = element_blank(),
        axis.title.x = element_text(face="bold", size=14),
        axis.title.y = element_text(face="bold", size=14),
        legend.title = element_text(size=12, face="bold"),
        legend.text = element_text(size=12),
        axis.ticks = element_blank(),
        plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm"), #top, right, bottom, left
        plot.title = element_text(size = 16, face="bold")) +
  guides(color = guide_legend(order=1, override.aes=list(size=3)),
         shape = guide_legend(order=2, override.aes=list(size=3))) 



### Color = Season, Shape = Lake_Type
plot_ordination(
  physeq = Ver.rel,
  ordination = nmds.bray,
  color = "Season",
  shape = "Lake_Type") + 
  scale_color_manual(values = c("indianred1","deepskyblue","mediumpurple2")) +
  scale_shape_manual(values = c(15,17,16)) +
  theme(axis.text.x = element_blank(), 
        axis.text.y = element_blank(),
        axis.title.x = element_text(face="bold", size=14),
        axis.title.y = element_text(face="bold", size=14),
        legend.title = element_text(size=12, face="bold"),
        legend.text = element_text(size=12),
        axis.ticks = element_blank(),
        plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm"), #top, right, bottom, left
        plot.title = element_text(size = 16, face="bold")) +
  guides(color = guide_legend(order=1, override.aes=list(size=3)),
         shape = guide_legend(order=2, override.aes=list(size=3))) 



### Facet = Lake_Type, Color = Fraction, Shape = Season
plot_ordination(
  physeq = Ver.rel,
  ordination = nmds.bray,
  color = "Fraction",
  shape = "Season") + 
  scale_color_manual(values = c("blue3","gold","tan3")) +
  facet_wrap(facets = ~Lake_Type) +
  theme(axis.text.x = element_blank(), 
        axis.text.y = element_blank(),
        axis.title.x = element_text(face="bold", size=14),
        axis.title.y = element_text(face="bold", size=14),
        legend.title = element_text(size=12, face="bold"),
        legend.text = element_text(size=12),
        axis.ticks = element_blank(),
        plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm"), #top, right, bottom, left
        plot.title = element_text(size = 16, face="bold")) +
  guides(color = guide_legend(order=1, override.aes=list(size=3)),
         shape = guide_legend(order=2, override.aes=list(size=3))) 
  


### Facet = Lake_Type, Color = Season, Shape = Fraction
bray.nmds.facet <- plot_ordination(
  physeq = Ver.rel,
  ordination = nmds.bray,
  color = "Season",
  shape = "Fraction") + 
  scale_color_manual(values = c("indianred1","deepskyblue","mediumpurple2")) +
  facet_wrap(facets = ~Lake_Type) +
  ggtitle("Bray-Curtis Dissimilarity") +
  theme(axis.text.x = element_blank(), 
        axis.text.y = element_blank(),
        axis.title.x = element_text(face="bold", size=14),
        axis.title.y = element_text(face="bold", size=14),
        legend.title = element_text(size=12, face="bold"),
        legend.text = element_text(size=12),
        axis.ticks = element_blank(),
        plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm"), #top, right, bottom, left
        plot.title = element_text(size = 16, face="bold")) +
  guides(color = guide_legend(order=1, override.aes=list(size=3)),
         shape = guide_legend(order=2, override.aes=list(size=3))) 

```



### NMDS (Sorensen)
``` {r soren}
###################################### SORENSEN NMDS ######################################
nowinOTU.df <- data.frame(otu_table(Ver.rel))
nmds.soren <- metaMDS(nowinOTU.df, distance="bray", binary=TRUE)
  # stress = 0.1934691 
nmds.soren <- data.frame(nmds.soren$points)
nmds.soren$names <- row.names(nmds.soren)
nowinSamp <- data.frame(sample_data(Ver.rel))
nmds.soren$Lake_Type <- nowinSamp$Lake_Type
nmds.soren$Lake <- nowinSamp$Lake
nmds.soren$Date <- nowinSamp$Date
nmds.soren$Site <- nowinSamp$Site
nmds.soren$Depth <- nowinSamp$Depth
nmds.soren$Fraction <- nowinSamp$Fraction
nmds.soren$Season <- nowinSamp$Season



### Color = Depth, Shape = Lake_Type
ggplot(nmds.soren, aes(MDS1*10, MDS2*10, color=Depth, shape = Lake_Type,)) +
  xlab("NMDS1") + ylab("NMDS2") + ggtitle("Sorensen Dissimilarity") +
  geom_point(size = 2, alpha = 1) + theme_bw() +
  scale_color_manual(name = "Depth", breaks = c("Top", "DCM", "Bottom", "Sediment"),
                     labels = c("Top", "DCM", "Bottom", "Sediment"),
                     values = c("indianred1","green3","mediumpurple2", "skyblue")) +
  scale_shape_manual(name = "Lake_Type", breaks = c("Laurentian","Estuary","Inland"),
                     labels = c("Laurentian","Estuary","Inland"),
                     values = c(15,17,16)) +
  theme(axis.text.x = element_blank(), 
        axis.text.y = element_blank(),
        axis.title.x = element_text(face="bold", size=14),
        axis.title.y = element_text(face="bold", size=14),
        legend.title = element_text(size=12, face="bold"),
        legend.text = element_text(size=12),
        axis.ticks = element_blank(),
        plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm"), #top, right, bottom, left
        plot.title = element_text(size = 16, face="bold")) +
  guides(color = guide_legend(order=1, override.aes=list(size=3)),
         shape = guide_legend(order=2, override.aes=list(size=3))) 


### Color = Lake_Type, Shape = Depth
soren.nmds.Lake_Type.depth <- ggplot(nmds.soren, aes(MDS1*10, MDS2*10, color=Lake_Type, shape = Depth,)) +
  xlab("NMDS1") + ylab("NMDS2") + ggtitle("Sorensen Dissimilarity") +
  geom_point(size = 2, alpha = 1) + theme_bw() +
  scale_color_manual(values = c("indianred1","deepskyblue","mediumpurple2")) +
  theme(axis.text.x = element_blank(), 
        axis.text.y = element_blank(),
        axis.title.x = element_text(face="bold", size=14),
        axis.title.y = element_text(face="bold", size=14),
        legend.title = element_text(size=12, face="bold"),
        legend.text = element_text(size=12),
        axis.ticks = element_blank(),
        plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm"), #top, right, bottom, left
        plot.title = element_text(size = 16, face="bold")) +
  guides(color = guide_legend(order=1, override.aes=list(size=3)),
         shape = guide_legend(order=2, override.aes=list(size=3))) 




### Color = Fraction, Shape = Depth
ggplot(nmds.soren, aes(MDS1*10, MDS2*10, color=Fraction, shape = Depth,)) +
  xlab("NMDS1") + ylab("NMDS2") + ggtitle("Sorensen Dissimilarity") +
  geom_point(size = 2, alpha = 1) + theme_bw() +
  scale_color_manual(values = c("indianred1","deepskyblue","mediumpurple2")) +
  scale_shape_manual(values = c(17, 5, 16, 4)) +
  theme(axis.text.x = element_blank(), 
        axis.text.y = element_blank(),
        axis.title.x = element_text(face="bold", size=14),
        axis.title.y = element_text(face="bold", size=14),
        legend.title = element_text(size=12, face="bold"),
        legend.text = element_text(size=12),
        axis.ticks = element_blank(),
        plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm"), #top, right, bottom, left
        plot.title = element_text(size = 16, face="bold")) +
  guides(color = guide_legend(order=1, override.aes=list(size=3)),
         shape = guide_legend(order=2, override.aes=list(size=3))) 


### Facet = Lake_Type, Color = Fraction, Shape = Season
ggplot(nmds.soren, aes(MDS1*10, MDS2*10, color=Fraction, shape = Season,)) +
  xlab("NMDS1") + ylab("NMDS2") + ggtitle("Sorensen Dissimilarity") +
  geom_point(size = 2, alpha = 1) + theme_bw() +
  facet_grid(.~Lake_Type) +
  scale_color_manual(values = c("blue3","gold", "tan3")) +
  scale_shape_manual(values = c(17, 16, 15)) +
  theme(axis.text.x = element_blank(), 
        axis.text.y = element_blank(),
        axis.title.x = element_text(face="bold", size=14),
        axis.title.y = element_text(face="bold", size=14),
        legend.title = element_text(size=12, face="bold"),
        legend.text = element_text(size=12),
        axis.ticks = element_blank(),
        plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm"), #top, right, bottom, left
        plot.title = element_text(size = 16, face="bold")) +
  guides(color = guide_legend(order=1, override.aes=list(size=3)),
         shape = guide_legend(order=2, override.aes=list(size=3)))



### Facet = Lake_Type, Color = Season, Shape = Fraction
soren.nmds.facet <- ggplot(nmds.soren, aes(MDS1*10, MDS2*10, color=Season, shape = Fraction,)) +
  xlab("NMDS1") + ylab("NMDS2") + ggtitle("Sorensen Dissimilarity") +
  geom_point(size = 2, alpha = 1) + theme_bw() +
  facet_grid(.~Lake_Type) +
  scale_color_manual(values = c("indianred1","deepskyblue","mediumpurple2")) +
  scale_shape_manual(values = c(16, 17, 15)) +
  theme(axis.text.x = element_blank(), 
        axis.text.y = element_blank(),
        axis.title.x = element_text(face="bold", size=14),
        axis.title.y = element_text(face="bold", size=14),
        legend.title = element_text(size=12, face="bold"),
        legend.text = element_text(size=12),
        axis.ticks = element_blank(),
        plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm"), #top, right, bottom, left
        plot.title = element_text(size = 16, face="bold")) +
  guides(color = guide_legend(order=1, override.aes=list(size=3)),
         shape = guide_legend(order=2, override.aes=list(size=3)))

```



### Create NMDS comparison plots
```{r compare.nmds}
#tiff("nmds.compare.tiff", width = 14, height = 10, units = "in", res = 500)
grid.newpage()
pushViewport(viewport(layout=(grid.layout(2,2, widths=c(1,1), heights=c(1.5,1.5)))))
print(bray.nmds.Lake_Type.depth, vp=viewport(layout.pos.row=1, layout.pos.col=1))
print(bray.nmds.facet, vp=viewport(layout.pos.row=1, layout.pos.col=2))
print(soren.nmds.Lake_Type.depth, vp=viewport(layout.pos.row=2, layout.pos.col=1))
print(soren.nmds.facet, vp=viewport(layout.pos.row=2, layout.pos.col=2))
#dev.off()

### Bray-Curtis seems to cluster a bit more clearly than Sorensen

```







### PCoA (Bray-Curtis)
```{r pcoa, echo=FALSE}
Ver.bray <- phyloseq::distance(physeq = Ver.rel, method = "bray")
Ver.pcoa <- ordinate(physeq = Ver.rel, method = "PCoA", distance = "bray")


#To determine approximately how much variation the axes should explain:
#Ver.pcoa$values
#look for broken stick value

### Depth
plot_ordination(
  physeq = Ver.rel, 
  ordination = Ver.pcoa, 
  color = "Depth",
  axes = c(1, 2),
  title = "Bray-Curtis Dissimilarity") +
  scale_color_manual(values = c("indianred1","deepskyblue","mediumpurple2")) +
  theme(axis.text.x = element_blank(), 
        axis.text.y = element_blank(),
        axis.title.x = element_text(face="bold", size=14),
        axis.title.y = element_text(face="bold", size=14),
        legend.title = element_text(size=12, face="bold"),
        legend.text = element_text(size=12),
        axis.ticks = element_blank(),
        plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm"), #top, right, bottom, left
        plot.title = element_text(size = 16, face="bold")) +
  guides(color = guide_legend(order=1, override.aes=list(size=3)),
         shape = guide_legend(order=2, override.aes=list(size=3)))

### Lake_Type
plot_ordination(
  physeq = Ver.rel, 
  ordination = Ver.pcoa, 
  color = "Lake_Type",
  axes = c(2,3),
  title = "Bray-Curtis Dissimilarity") +
  scale_color_manual(values = c("indianred1","deepskyblue","mediumpurple2")) +
  theme(axis.text.x = element_blank(), 
        axis.text.y = element_blank(),
        axis.title.x = element_text(face="bold", size=14),
        axis.title.y = element_text(face="bold", size=14),
        legend.title = element_text(size=12, face="bold"),
        legend.text = element_text(size=12),
        axis.ticks = element_blank(),
        plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm"), #top, right, bottom, left
        plot.title = element_text(size = 16, face="bold")) +
  guides(color = guide_legend(order=1, override.aes=list(size=3)),
         shape = guide_legend(order=2, override.aes=list(size=3)))

### Color = Lake_Type, Shape = Depth
bray.pcoa.Lake_Type.depth <- plot_ordination(
  physeq = Ver.rel, 
  ordination = Ver.pcoa, 
  color = "Lake_Type",
  shape = "Depth",
  axes = c(2,3),
  title = "Bray-Curtis Dissimilarity") +
  scale_color_manual(values = c("indianred1","deepskyblue","mediumpurple2")) +
  theme(axis.text.x = element_blank(), 
        axis.text.y = element_blank(),
        axis.title.x = element_text(face="bold", size=14),
        axis.title.y = element_text(face="bold", size=14),
        legend.title = element_text(size=12, face="bold"),
        legend.text = element_text(size=12),
        axis.ticks = element_blank(),
        plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm"), #top, right, bottom, left
        plot.title = element_text(size = 16, face="bold")) +
  guides(color = guide_legend(order=1, override.aes=list(size=3)),
         shape = guide_legend(order=2, override.aes=list(size=3)))

# Fraction PCoA
plot_ordination(
  physeq = Ver.rel, 
  ordination = Ver.pcoa, 
  color = "Fraction",
  shape = "Season",
  axes = c(1, 3),
  title = "Bray-Curtis Dissimilarity") +
  facet_wrap(facets = ~Lake_Type) +
  theme(axis.text.x = element_blank(), 
        axis.text.y = element_blank(),
        axis.title.x = element_text(face="bold", size=14),
        axis.title.y = element_text(face="bold", size=14),
        legend.title = element_text(size=12, face="bold"),
        legend.text = element_text(size=12),
        axis.ticks = element_blank(),
        plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm"), #top, right, bottom, left
        plot.title = element_text(size = 16, face="bold")) +
  guides(color = guide_legend(order=1, override.aes=list(size=3)),
         shape = guide_legend(order=2, override.aes=list(size=3)))

# Season PCoA
bray.pcoa.facet <- plot_ordination(
  physeq = Ver.rel, 
  ordination = Ver.pcoa, 
  color = "Season",
  shape = "Fraction",
  axes = c(1,2),
  title = "Bray-Curtis Dissimilarity") +
  facet_wrap(facets = ~Lake_Type) +
  scale_color_manual(values = c("indianred1","deepskyblue","mediumpurple2")) +
  theme(axis.text.x = element_blank(), 
        axis.text.y = element_blank(),
        axis.title.x = element_text(face="bold", size=14),
        axis.title.y = element_text(face="bold", size=14),
        legend.title = element_text(size=12, face="bold"),
        legend.text = element_text(size=12),
        axis.ticks = element_blank(),
        plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm"), #top, right, bottom, left
        plot.title = element_text(size = 16, face="bold")) +
  guides(color = guide_legend(order=1, override.aes=list(size=3)),
         shape = guide_legend(order=2, override.aes=list(size=3)))
```



### PCoA (Sorensen)
``` {r pcoa.sorensen}
test <- dsvdis(nowinOTU.df, index="sorensen")
soren.pcoa<- ordinate(physeq = Ver.rel, method = "PCoA", distance = test)

### Depth
plot_ordination(
  physeq = Ver.rel, 
  ordination = soren.pcoa, 
  color = "Depth",
  axes = c(1,2),
  title = "Sorensen Dissimilarity") +
  theme(axis.text.x = element_blank(), 
        axis.text.y = element_blank(),
        axis.title.x = element_text(face="bold", size=14),
        axis.title.y = element_text(face="bold", size=14),
        legend.title = element_text(size=12, face="bold"),
        legend.text = element_text(size=12),
        axis.ticks = element_blank(),
        plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm"), #top, right, bottom, left
        plot.title = element_text(size = 16, face="bold"))


### Lake_Type
plot_ordination(
  physeq = Ver.rel, 
  ordination = soren.pcoa, 
  color = "Lake_Type",
  axes = c(2,3),
  title = "Sorensen Dissimilarity") +
  theme(axis.text.x = element_blank(), 
        axis.text.y = element_blank(),
        axis.title.x = element_text(face="bold", size=14),
        axis.title.y = element_text(face="bold", size=14),
        legend.title = element_text(size=12, face="bold"),
        legend.text = element_text(size=12),
        axis.ticks = element_blank(),
        plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm"), #top, right, bottom, left
        plot.title = element_text(size = 16, face="bold"))


### Color = Lake_Type, Shape = Depth
soren.pcoa.Lake_Type.depth <- plot_ordination(
  physeq = Ver.rel, 
  ordination = soren.pcoa, 
  color = "Lake_Type",
  shape = "Depth",
  axes = c(2,3),
  title = "Sorensen Dissimilarity") +
  scale_color_manual(values = c("indianred1","deepskyblue","mediumpurple2")) +
  theme(axis.text.x = element_blank(), 
        axis.text.y = element_blank(),
        axis.title.x = element_text(face="bold", size=14),
        axis.title.y = element_text(face="bold", size=14),
        legend.title = element_text(size=12, face="bold"),
        legend.text = element_text(size=12),
        axis.ticks = element_blank(),
        plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm"), #top, right, bottom, left
        plot.title = element_text(size = 16, face="bold")) +
  guides(color = guide_legend(order=1, override.aes=list(size=3)),
         shape = guide_legend(order=2, override.aes=list(size=3)))


### Facet = Lake_Type, Color = Fraction, Shape = Season
plot_ordination(
  physeq = Ver.rel, 
  ordination = soren.pcoa, 
  color = "Fraction",
  shape = "Season",
  axes = c(1, 3),
  title = "Sorensen Dissimilarity") +
  facet_wrap(facets = ~Lake_Type) +
  theme(axis.text.x = element_blank(), 
        axis.text.y = element_blank(),
        axis.title.x = element_text(face="bold", size=14),
        axis.title.y = element_text(face="bold", size=14),
        legend.title = element_text(size=12, face="bold"),
        legend.text = element_text(size=12),
        axis.ticks = element_blank(),
        plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm"), #top, right, bottom, left
        plot.title = element_text(size = 16, face="bold")) +
  guides(color = guide_legend(order=1, override.aes=list(size=3)),
         shape = guide_legend(order=2, override.aes=list(size=3)))

### Facet = Lake_Type, Color = Season, Shape = Fraction
soren.pcoa.facet <- plot_ordination(
  physeq = Ver.rel, 
  ordination = Ver.pcoa, 
  color = "Season",
  shape = "Fraction",
  axes = c(1,2),
  title = "Sorensen Dissimilarity") +
  facet_wrap(facets = ~Lake_Type) +
  scale_color_manual(values = c("indianred1","deepskyblue","mediumpurple2")) +
  theme(axis.text.x = element_blank(), 
        axis.text.y = element_blank(),
        axis.title.x = element_text(face="bold", size=14),
        axis.title.y = element_text(face="bold", size=14),
        legend.title = element_text(size=12, face="bold"),
        legend.text = element_text(size=12),
        axis.ticks = element_blank(),
        plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm"), #top, right, bottom, left
        plot.title = element_text(size = 16, face="bold")) +
  guides(color = guide_legend(order=1, override.aes=list(size=3)),
         shape = guide_legend(order=2, override.aes=list(size=3)))

```


### Create PCoA comparison plots
```{r compare.pcoa}
tiff("pcoa.compare2.tiff", width = 14, height = 10, units = "in", res = 500)
grid.newpage()
pushViewport(viewport(layout=(grid.layout(2,2, widths=c(1,1), heights=c(1.5,1.5)))))
print(bray.pcoa.Lake_Type.depth, vp=viewport(layout.pos.row=1, layout.pos.col=1))
print(bray.pcoa.facet, vp=viewport(layout.pos.row=1, layout.pos.col=2))
print(soren.pcoa.Lake_Type.depth, vp=viewport(layout.pos.row=2, layout.pos.col=1))
print(soren.pcoa.facet, vp=viewport(layout.pos.row=2, layout.pos.col=2))
dev.off()

### Bray-Curtis seems to cluster a bit more clearly than Sorensen

```



### Ver Lake Type PCoA
```{r ver.type.pcoa}
### Laurentian
Ver.lau <- subset_samples(Ver.rel, Lake_Type == "Laurentian")
Ver.lau.bray <- phyloseq::distance(physeq = Ver.lau, method = "bray")
Ver.lau.pcoa <- ordinate(physeq = Ver.lau, method = "PCoA", distance = "bray")


Ver.lau.ord <- plot_ordination(physeq = Ver.lau, ordination = Ver.lau.pcoa, axes = c(1, 2),
                title= "Bray-Curtis Dissimilarity") + 
  geom_point(size=2, aes(color = Season, shape = Depth,
                         fill = factor(ifelse(Fraction == "Free", NA, Season)),
                         alpha = Fraction)) +
  scale_shape_manual(values = c(21, 24, 22)) +
  scale_fill_discrete(na.value = "white", guide = "none") +
  scale_alpha_manual(values=c("Particle"=1, "Free"=1)) +
  theme(axis.text.x = element_blank(), 
        axis.text.y = element_blank(),
        axis.title.x = element_text(face="bold", size=10),
        axis.title.y = element_text(face="bold", size=10),
        legend.title = element_text(size=8, face="bold"),
        legend.text = element_text(size=8),
        axis.ticks = element_blank(),
        plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm"), #top, right, bottom, left
        plot.title = element_text(size = 10, face="bold")) +
  annotate("text", x= -0.38, y=0.45, label = "A", fontface = "bold", size=6, colour="black") +
  guides(colour=FALSE, shape=FALSE, alpha = FALSE)
  #guides(color = guide_legend(order=1, title="Season", override.aes=list(size=3)),
  #       shape = guide_legend(order=2, title="Depth",override.aes=list(size=3)),
  #       alpha = guide_legend(order=3, title="Fraction",override.aes=list(size=3))) 




  
# Particle = fill, Free = empty
# Note: for scale_fill_discrete to work, you MUST set scale_shape_manual to between 21-25
#        You also can't set colors manually



Ver.est <- subset_samples(Ver.rel, Lake_Type == "Estuary")
Ver.est.bray <- phyloseq::distance(physeq = Ver.est, method = "bray")
Ver.est.pcoa <- ordinate(physeq = Ver.est, method = "PCoA", distance = "bray")
  
Ver.est.ord <- plot_ordination(physeq = Ver.est, ordination = Ver.est.pcoa, axes = c(1, 2),
                title= "Bray-Curtis Dissimilarity") + 
  geom_point(size=2, aes(color = Season, shape = Depth,
                         fill = factor(ifelse(Fraction == "Free", NA, Season)),
                         alpha = Fraction)) +
  scale_shape_manual(values = c(21, 24, 22)) +
  scale_fill_discrete(na.value = "white", guide = "none") +
  scale_alpha_manual(values=c("Particle"=1, "Free"=1, "Sediment"=0.5)) +
  theme(axis.text.x = element_blank(), 
        axis.text.y = element_blank(),
        axis.title.x = element_text(face="bold", size=10),
        axis.title.y = element_text(face="bold", size=10),
        legend.title = element_text(size=8, face="bold"),
        legend.text = element_text(size=8),
        axis.ticks = element_blank(),
        plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm"), #top, right, bottom, left
        plot.title = element_text(size = 10, face="bold")) +
  annotate("text", x= -0.29, y=0.45, label = "B", fontface = "bold", size=6, colour="black") +
  guides(colour=FALSE, shape=FALSE, alpha = FALSE) 
  #guides(color = guide_legend(order=1, title="Season", override.aes=list(size=3)),
  #       shape = guide_legend(order=2, title="Depth",override.aes=list(size=3)),
  #       alpha = guide_legend(order=3, title="Fraction",override.aes=list(size=3)))


 ##### Let's steal the legend 
 #g_legend <- function(a.gplot){
#   tmp <- ggplot_gtable(ggplot_build(a.gplot))
#   leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
#   legend <- tmp$grobs[[leg]]
#   return(legend)}
 legend <- g_legend(Ver.est.ord)

  
  
Ver.inl <- subset_samples(Ver.rel, Lake_Type == "Inland")
Ver.inl.bray <- phyloseq::distance(physeq = Ver.inl, method = "bray")
Ver.inl.pcoa <- ordinate(physeq = Ver.inl, method = "PCoA", distance = "bray")

Ver.inl.ord <- plot_ordination(physeq = Ver.inl, ordination = Ver.inl.pcoa, axes = c(1, 2),
                title= "Bray-Curtis Dissimilarity") + 
  geom_point(size=2, aes(color = Season, shape = Depth,
                         fill = factor(ifelse(Fraction == "Free", NA, Season)),
                         alpha = Fraction)) +
  scale_shape_manual(values = c(21, 24, 22)) +
  scale_fill_discrete(na.value = "white", guide = "none") +
  scale_alpha_manual(values=c("Particle"=1, "Free"=1)) +
  theme(axis.text.x = element_blank(), 
        axis.text.y = element_blank(),
        axis.title.x = element_text(face="bold", size=10),
        axis.title.y = element_text(face="bold", size=10),
        legend.title = element_text(size=8, face="bold"),
        legend.text = element_text(size=8),
        axis.ticks = element_blank(),
        plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm"), #top, right, bottom, left
        plot.title = element_text(size = 10, face="bold")) +
  annotate("text", x= -0.46, y=0.69, label = "C", fontface = "bold", size=6, colour="black") +
  guides(colour=FALSE, shape=FALSE, alpha = FALSE) 
  #guides(color = guide_legend(order=1, title="Season", override.aes=list(size=3)),
  #       shape = guide_legend(order=2, title="Depth",override.aes=list(size=3)),
  #       alpha = guide_legend(order=3, title="Fraction",override.aes=list(size=3)))




#tiff("ver.pcoa.tiff", width = 15, height = 10, units = "cm", res = 500)
#grid.newpage()
#pushViewport(viewport(layout=(grid.layout(1,3, widths=c(1,1,1), heights=c(1,1,1)))))
#print(Ver.lau.ord, vp=viewport(layout.pos.row=1, layout.pos.col=1))
#print(Ver.est.ord, vp=viewport(layout.pos.row=1, layout.pos.col=2))
#print(Ver.inl.ord, vp=viewport(layout.pos.row=1, layout.pos.col=3))
#dev.off()



```

### PERMANOVA/Adonis
```{r permanova, echo=FALSE}
sampledf <- data.frame(sample_data(Ver.rel))

##### Depth #####
adonis(Ver.bray ~ Depth, data=sampledf)
  ## R2 = 0.08797
  ## p-value = 0.001
beta.tax = betadisper(d=Ver.bray, group=sampledf$Depth)
p <- permutest(beta.tax)
p$tab
  # p-value = 0.001
  # F = 28.00798

# Top vs. Bottom
adonis.tb <- subset_samples(physeq = Ver.rel, Depth != "DCM")
adonis.tb <- subset_samples(physeq = adonis.tb, Depth != "Sediment")
adonis.tb.bray <- phyloseq::distance(physeq = adonis.tb, method = "bray")
adonis.tb <- data.frame(sample_data(adonis.tb))
adonis(adonis.tb.bray ~ Depth, data=adonis.tb)
  # p-value = 0.004
  # R2 = 0.01411
beta.tax = betadisper(d=adonis.tb.bray, group=adonis.tb$Depth)
p <- permutest(beta.tax)
p$tab
  # p-value = 0.498
  # F = 0.5144987

# Top vs. Sediment
adonis.ts <- subset_samples(physeq = Ver.rel, Depth != "DCM")
adonis.ts <- subset_samples(physeq = adonis.ts, Depth != "Bottom")
adonis.ts.bray <- phyloseq::distance(physeq = adonis.ts, method = "bray")
adonis.ts <- data.frame(sample_data(adonis.ts))
adonis(adonis.ts.bray ~ Depth, data=adonis.ts)
  # p-value = 0.001
  # R2 = 0.13396
beta.tax = betadisper(d=adonis.ts.bray, group=adonis.ts$Depth)
p <- permutest(beta.tax)
p$tab
  # p-value = 0.001
  # F = 58.66768

# Bottom vs. Sediment
adonis.bs <- subset_samples(physeq = Ver.rel, Depth != "DCM")
adonis.bs <- subset_samples(physeq = adonis.bs, Depth != "Top")
adonis.bs.bray <- phyloseq::distance(physeq = adonis.bs, method = "bray")
adonis.bs <- data.frame(sample_data(adonis.bs))
adonis(adonis.bs.bray ~ Depth, data=adonis.bs)
  # p-value = 0.001
  # R2 = 0.13453
beta.tax = betadisper(d=adonis.bs.bray, group=adonis.bs$Depth)
p <- permutest(beta.tax)
p$tab
  # p-value = 0.001
  # F = 47.73623


##### Lake_Type #####
adonis(Ver.bray ~ Lake_Type, data=sampledf)
  ## R2 = 0.10932
  ## p-value = 0.001
beta.tax = betadisper(Ver.bray, sampledf$Lake_Type)
p <- permutest(beta.tax)
p$tab
  ## p-value = 0.77
  ## F = 0.2776733

# Laurentian vs. Estuary
adonis.scaled.le <- subset_samples(physeq = scaled, Lake_Type != "Inland")
adonis.scaled.le.bray <- phyloseq::distance(physeq = adonis.scaled.le, method = "bray")
adonis.scaled.le <- data.frame(sample_data(adonis.scaled.le))
adonis(adonis.scaled.le.bray ~ Lake_Type, data=adonis.scaled.le)
  # p-value = 0.001
  # R2 = 0.09594 
beta.tax = betadisper(d=adonis.scaled.le.bray, group=adonis.scaled.le$Lake_Type)
p <- permutest(beta.tax)
p$tab
  # p-value = 0.092
  # F = 3.061581

# Laurentian vs. Inland
adonis.scaled.li <- subset_samples(physeq = scaled, Lake_Type != "Estuary")
adonis.scaled.li.bray <- phyloseq::distance(physeq = adonis.scaled.li, method = "bray")
adonis.scaled.li <- data.frame(sample_data(adonis.scaled.li))
adonis(adonis.scaled.li.bray ~ Lake_Type, data=adonis.scaled.li)
  # p-value = 0.001
  # R2 = 0.0732
beta.tax = betadisper(d=adonis.scaled.li.bray, group=adonis.scaled.li$Lake_Type)
p <- permutest(beta.tax)
p$tab
  # p-value = 0.001
  # F = 13.455

# Estuary vs. Inland
adonis.scaled.ei <- subset_samples(physeq = scaled, Lake_Type != "Laurentian")
adonis.scaled.ei.bray <- phyloseq::distance(physeq = adonis.scaled.ei, method = "bray")
adonis.scaled.ei <- data.frame(sample_data(adonis.scaled.ei))
adonis(adonis.scaled.ei.bray ~ Lake_Type, data=adonis.scaled.ei)
  # p-value = 0.001
  # R2 = 0.06387
beta.tax = betadisper(d=adonis.scaled.ei.bray, group=adonis.scaled.ei$Lake_Type)
p <- permutest(beta.tax)
p$tab
  # p-value = 0.256
  # F = 1.280102





##### Fraction #####
adonis(Ver.bray ~ Fraction, data=sampledf)
  ## p-value = 0.001
  ## R2 = 0.16121
beta.tax = betadisper(Ver.bray, sampledf$Fraction)
p <- permutest(beta.tax)
p$tab
  ## P-value = 0.001
  ## F = 29.83702

# Particle vs. Free
adonis.pf <- subset_samples(physeq = Ver.rel, Fraction != "Sediment")
adonis.pf.bray <- phyloseq::distance(physeq = adonis.pf, method = "bray")
adonis.pf <- data.frame(sample_data(adonis.pf))
adonis(adonis.pf.bray ~ Fraction, data=adonis.pf)
  # p-value = 0.001
  # R2 = 0.09474
beta.tax = betadisper(d=adonis.pf.bray, group=adonis.pf$Fraction)
p <- permutest(beta.tax)
p$tab
  # p-value = 0.001
  # F = 23.66752

# Particle vs. Sediment
adonis.ps <- subset_samples(physeq = Ver.rel, Fraction != "Free")
adonis.ps.bray <- phyloseq::distance(physeq = adonis.ps, method = "bray")
adonis.ps <- data.frame(sample_data(adonis.ps))
adonis(adonis.ps.bray ~ Fraction, data=adonis.ps)
  # p-value = 0.001
  # R2 = 0.12536
beta.tax = betadisper(d=adonis.ps.bray, group=adonis.ps$Fraction)
p <- permutest(beta.tax)
p$tab
  # p-value = 0.001
  # F = 58.6663

# Free vs. Sediment
adonis.fs <- subset_samples(physeq = Ver.rel, Fraction != "Particle")
adonis.fs.bray <- phyloseq::distance(physeq = adonis.fs, method = "bray")
adonis.fs <- data.frame(sample_data(adonis.fs))
adonis(adonis.fs.bray ~ Fraction, data=adonis.fs)
  # p-value = 0.001
  # R2 = 0.17969
beta.tax = betadisper(d=adonis.fs.bray, group=adonis.fs$Fraction)
p <- permutest(beta.tax)
p$tab
  # p-value = 0.001
  # F = 19.22966



##### Season #####
adonis(Ver.bray ~ Season, data=sampledf)
  ## p-value = 0.001
  ## R2 = 0.0952
beta.tax = betadisper(Ver.bray, sampledf$Season)
p <- permutest(beta.tax)
p$tab
  ## P-value = 0.008
  ## F = 4.999286

# Spring vs. Summer
adonis.ss <- subset_samples(physeq = Ver.rel, Season != "Fall")
adonis.ss.bray <- phyloseq::distance(physeq = adonis.ss, method = "bray")
adonis.ss <- data.frame(sample_data(adonis.ss))
adonis(adonis.ss.bray ~ Season, data=adonis.ss)
  # p-value = 0.001
  # R2 = 0.09052
beta.tax = betadisper(d=adonis.ss.bray, group=adonis.ss$Fraction)
p <- permutest(beta.tax)
p$tab
  # p-value = 0.001
  # F = 18.14419

# Spring vs. Fall
adonis.sf <- subset_samples(physeq = Ver.rel, Season != "Summer")
adonis.sf.bray <- phyloseq::distance(physeq = adonis.sf, method = "bray")
adonis.sf <- data.frame(sample_data(adonis.sf))
adonis(adonis.sf.bray ~ Season, data=adonis.sf)
  # p-value = 0.001
  # R2 = 0.13916
beta.tax = betadisper(d=adonis.sf.bray, group=adonis.sf$Fraction)
p <- permutest(beta.tax)
p$tab
  # p-value = 0.001
  # F = 22.43885

# Summer vs. Fall
adonis.suf <- subset_samples(physeq = Ver.rel, Season != "Spring")
adonis.suf.bray <- phyloseq::distance(physeq = adonis.suf, method = "bray")
adonis.suf <- data.frame(sample_data(adonis.suf))
adonis(adonis.suf.bray ~ Season, data=adonis.suf)
  # p-value = 0.001
  # R2 = 0.02691
beta.tax = betadisper(d=adonis.suf.bray, group=adonis.suf$Fraction)
p <- permutest(beta.tax)
p$tab
  # p-value = 0.001
  # F = 20.66774


##### Nested adonis #####
adonis(Ver.bray ~ Lake_Type + Depth + Fraction + Season, data = sampledf)
  ## Lake_Type R2 = 0.10932
  ## Depth R2 = 0.09008
  ## Fraction R2 = 0.08552
  ## Season R2 = 0.07912
  ## all p-values = 0.001



```




### Ver Lake PERMANOVA
```{r v.lake.adonis}

########## LAURENTIAN ##########
sampledf <- data.frame(sample_data(Ver.lau))

### Depth
adonis(Ver.lau.bray ~ Depth, data=sampledf)
  # R2 = 0.04999
  # p-value = 0.119
beta.tax = betadisper(d=Ver.lau.bray, group=sampledf$Depth)
p <- permutest(beta.tax)
p$tab
  # p-value = 0.028
  # F = 4.689563

### Season
adonis(Ver.lau.bray ~ Season, data=sampledf)
  # R2 = 0.43082
  # p-value = 0.001
beta.tax = betadisper(d=Ver.lau.bray, group=sampledf$Season)
p <- permutest(beta.tax)
p$tab
  # p-value = 0.001
  # F = 11.22

# Spring vs. Summer
adonis.ss <- subset_samples(physeq = Ver.lau, Season != "Fall")
adonis.ss.bray <- phyloseq::distance(physeq = adonis.ss, method = "bray")
adonis.ss <- data.frame(sample_data(adonis.ss))
adonis(adonis.ss.bray ~ Season, data=adonis.ss)
  # R2 = 0.47247
  # p-value = 0.001
beta.tax = betadisper(d=adonis.ss.bray, group=adonis.ss$Season)
p <- permutest(beta.tax)
p$tab
  # p-value = 0.001
  # F = 15.00923

# Spring vs. Fall
adonis.spf <- subset_samples(physeq = Ver.lau, Season != "Summer")
adonis.spf.bray <- phyloseq::distance(physeq = adonis.spf, method = "bray")
adonis.spf <- data.frame(sample_data(adonis.spf))
adonis(adonis.spf.bray ~ Season, data=adonis.spf)
  # R2 = 0.41549
  # p-value = 0.001
beta.tax = betadisper(d=adonis.spf.bray, group=adonis.spf$Season)
p <- permutest(beta.tax)
p$tab
  # p-value = 0.001
  # F = 15.21934

# Summer vs. Fall
adonis.suf <- subset_samples(physeq = Ver.lau, Season != "Spring")
adonis.suf.bray <- phyloseq::distance(physeq = adonis.suf, method = "bray")
adonis.suf <- data.frame(sample_data(adonis.suf))
adonis(adonis.suf.bray ~ Season, data=adonis.suf)
  # R2 = 0.18953
  # p-value = 0.001
beta.tax = betadisper(d=adonis.suf.bray, group=adonis.suf$Season)
p <- permutest(beta.tax)
p$tab
  # p-value = 0.825
  # F = 0.04694171

### Fraction
adonis(Ver.lau.bray ~ Fraction, data=sampledf)
  # R2 = 0.17341
  # p-value = 0.001
beta.tax = betadisper(d=Ver.lau.bray, group=sampledf$Fraction)
p <- permutest(beta.tax)
p$tab
  # p-value = 0.706
  # F = 0.1308235 

##### Nested adonis #####
adonis(Ver.lau.bray ~ Season + Fraction + Depth, data = sampledf)
  ## Season R2= 0.43082; p-value = 0.001
  ## Fraction R2= 0.17534; p-value = 0.001
  ## Depth R2= 0.04616; p-value = 0.007


########## ESTUARY ##########
sampledf <- data.frame(sample_data(Ver.est))

### Depth
adonis(Ver.est.bray ~ Depth, data=sampledf)
  # R2 = 0.30094
  # p-value = 0.001
beta.tax = betadisper(d=Ver.est.bray, group=sampledf$Depth)
p <- permutest(beta.tax)
p$tab
  # p-value = 0.019
  # F = 4.068133

# Water vs. Sediment
adonis(Ver.est.bray ~ Source, data=sampledf)
  # R2 = 0.28633
  # p-value = 0.001
beta.tax = betadisper(d=Ver.est.bray, group=sampledf$Source)
p <- permutest(beta.tax)
p$tab
  # p-value = 0.004
  # F = 9.392686

# Top vs. Bottom
adonis.tb <- subset_samples(physeq = Ver.est, Depth != "Sediment")
adonis.tb.bray <- phyloseq::distance(physeq = adonis.tb, method = "bray")
adonis.tb <- data.frame(sample_data(adonis.tb))
adonis(adonis.tb.bray ~ Depth, data=adonis.tb)
  # R2 = 0.0224
  # p-value = 0.313
beta.tax = betadisper(d=adonis.tb.bray, group=adonis.tb$Depth)
p <- permutest(beta.tax)
p$tab
  # p-value = 0.811
  # F = 0.05359156

# Top vs. Sediment
adonis.ts <- subset_samples(physeq = Ver.est, Depth != "Bottom")
adonis.ts.bray <- phyloseq::distance(physeq = adonis.ts, method = "bray")
adonis.ts <- data.frame(sample_data(adonis.ts))
adonis(adonis.ts.bray ~ Depth, data=adonis.ts)
  # R2 = 0.39375
  # p-value = 0.001
beta.tax = betadisper(d=adonis.ts.bray, group=adonis.ts$Depth)
p <- permutest(beta.tax)
p$tab
  # p-value = 0.016
  # F = 6.536549

# Bottom vs. Sediment
adonis.bs <- subset_samples(physeq = Ver.est, Depth != "Top")
adonis.bs.bray <- phyloseq::distance(physeq = adonis.bs, method = "bray")
adonis.bs <- data.frame(sample_data(adonis.bs))
adonis(adonis.bs.bray ~ Depth, data=adonis.bs)
  # R2 = 0.39372
  # p-value = 0.001
beta.tax = betadisper(d=adonis.bs.bray, group=adonis.bs$Depth)
p <- permutest(beta.tax)
p$tab
  # p-value = 0.009
  # F = 8.668537

### Season
adonis(Ver.est.bray ~ Season, data=sampledf)
  # R2 = 0.21914
  # p-value = 0.001
beta.tax = betadisper(d=Ver.est.bray, group=sampledf$Season)
p <- permutest(beta.tax)
p$tab
  # p-value = 0.756
  # F = 0.2979526

# Spring vs. Summer
adonis.ss <- subset_samples(physeq = Ver.est, Season != "Fall")
adonis.ss.bray <- phyloseq::distance(physeq = adonis.ss, method = "bray")
adonis.ss <- data.frame(sample_data(adonis.ss))
adonis(adonis.ss.bray ~ Season, data=adonis.ss)
  # R2 = 0.21356
  # p-value = 0.001
beta.tax = betadisper(d=adonis.ss.bray, group=adonis.ss$Season)
p <- permutest(beta.tax)
p$tab
  # p-value = 0.814
  # F = 0.05572213

# Spring vs. Fall
adonis.spf <- subset_samples(physeq = Ver.est, Season != "Summer")
adonis.spf.bray <- phyloseq::distance(physeq = adonis.spf, method = "bray")
adonis.spf <- data.frame(sample_data(adonis.spf))
adonis(adonis.spf.bray ~ Season, data=adonis.spf)
  # R2 = 0.25881
  # p-value = 0.001
beta.tax = betadisper(d=adonis.spf.bray, group=adonis.spf$Season)
p <- permutest(beta.tax)
p$tab
  # p-value = 0.472
  # F = 0.5064546 

# Summer vs. Fall
adonis.suf <- subset_samples(physeq = Ver.est, Season != "Spring")
adonis.suf.bray <- phyloseq::distance(physeq = adonis.suf, method = "bray")
adonis.suf <- data.frame(sample_data(adonis.suf))
adonis(adonis.suf.bray ~ Season, data=adonis.suf)
  # R2 = 0.05553
  # p-value = 0.026
beta.tax = betadisper(d=adonis.suf.bray, group=adonis.suf$Season)
p <- permutest(beta.tax)
p$tab
  # p-value = 0.588
  # F = 0.2926064

### Fraction
adonis(Ver.est.bray ~ Fraction, data=sampledf)
  # R2 = 0.4132
  # p-value = 0.001
beta.tax = betadisper(d=Ver.est.bray, group=sampledf$Fraction)
p <- permutest(beta.tax)
p$tab
  # p-value = 0.086
  # F = 2.717009

# Particle vs. Free
adonis.frac <- subset_samples(physeq = Ver.est, Fraction != "Sediment")
adonis.frac.bray <- phyloseq::distance(physeq = adonis.frac, method = "bray")
adonis.frac <- data.frame(sample_data(adonis.frac))
adonis(adonis.frac.bray ~ Fraction, data=adonis.frac)
  # R2 = 0.19446
  # p-value = 0.001
beta.tax = betadisper(d=adonis.frac.bray, group=adonis.frac$Fraction)
p <- permutest(beta.tax)
p$tab
  # p-value = 0.141
  # F = 2.431074

# Particle vs. Sediment
adonis.ps <- subset_samples(physeq = Ver.est, Fraction != "Free")
adonis.ps.bray <- phyloseq::distance(physeq = adonis.ps, method = "bray")
adonis.ps <- data.frame(sample_data(adonis.ps))
adonis(adonis.ps.bray ~ Fraction, data=adonis.ps)
  # R2 = 0.39181
  # p-value = 0.001
beta.tax = betadisper(d=adonis.ps.bray, group=adonis.ps$Fraction)
p <- permutest(beta.tax)
p$tab
  # p-value = 0.076
  # F = 3.423727

# Free vs. Sediment
adonis.fs <- subset_samples(physeq = Ver.est, Fraction != "Particle")
adonis.fs.bray <- phyloseq::distance(physeq = adonis.fs, method = "bray")
adonis.fs <- data.frame(sample_data(adonis.fs))
adonis(adonis.fs.bray ~ Fraction, data=adonis.fs)
  # R2 = 0.51146
  # p-value = 0.001
beta.tax = betadisper(d=adonis.fs.bray, group=adonis.fs$Fraction)
p <- permutest(beta.tax)
p$tab
  # p-value = 0.235
  # F = 1.431922


##### Nested adonis #####
adonis(Ver.est.bray ~ Depth + Season + Fraction, data = sampledf)
  ## Depth R2= 0.30094; p-value = 0.001
  ## Season R2= 0.21645; p-value = 0.001
  ## Fraction R2= 0.12687; p-value = 0.001




########## INLAND ##########
sampledf <- data.frame(sample_data(Ver.inl))

### Depth
adonis(Ver.inl.bray ~ Depth, data=sampledf)
  # R2 = 0.03548
  # p-value = 0.002
beta.tax = betadisper(d=Ver.inl.bray, group=sampledf$Depth)
p <- permutest(beta.tax)
p$tab
  # p-value = 0.003
  # F = 9.534813

### Season
adonis(Ver.inl.bray ~ Season, data=sampledf)
  # R2 = 0.7723
  # p-value = 0.001
beta.tax = betadisper(d=Ver.inl.bray, group=sampledf$Season)
p <- permutest(beta.tax)
p$tab
  # p-value = 0.002
  # F = 8.038783

# Spring vs. Summer
adonis.ss <- subset_samples(physeq = Ver.inl, Season != "Fall")
adonis.ss.bray <- phyloseq::distance(physeq = adonis.ss, method = "bray")
adonis.ss <- data.frame(sample_data(adonis.ss))
adonis(adonis.ss.bray ~ Season, data=adonis.ss)
  # R2 = 0.06701
  # p-value = 0.001
beta.tax = betadisper(d=adonis.ss.bray, group=adonis.ss$Season)
p <- permutest(beta.tax)
p$tab
  # p-value = 0.001
  # F = 16.4969

# Spring vs. Fall
adonis.spf <- subset_samples(physeq = Ver.inl, Season != "Summer")
adonis.spf.bray <- phyloseq::distance(physeq = adonis.spf, method = "bray")
adonis.spf <- data.frame(sample_data(adonis.spf))
adonis(adonis.spf.bray ~ Season, data=adonis.spf)
  # R2 = 0.17293
  # p-value = 0.001
beta.tax = betadisper(d=adonis.spf.bray, group=adonis.spf$Season)
p <- permutest(beta.tax)
p$tab
  # p-value = 0.42
  # F = 0.7247857

# Summer vs. Fall
adonis.suf <- subset_samples(physeq = Ver.inl, Season != "Spring")
adonis.suf.bray <- phyloseq::distance(physeq = adonis.suf, method = "bray")
adonis.suf <- data.frame(sample_data(adonis.suf))
adonis(adonis.suf.bray ~ Season, data=adonis.suf)
  # R2 = 0.02092
  # p-value = 0.019
beta.tax = betadisper(d=adonis.suf.bray, group=adonis.suf$Season)
p <- permutest(beta.tax)
p$tab
  # p-value = 0.042
  # F = 4.181126

### Fraction
adonis(Ver.inl.bray ~ Fraction, data=sampledf)
  # R2 = 0.11226
  # p-value = 0.001
beta.tax = betadisper(d=Ver.inl.bray, group=sampledf$Fraction)
p <- permutest(beta.tax)
p$tab
  # p-value = 0.006
  # F = 8.083098

##### Nested adonis #####
adonis(Ver.inl.bray ~ Fraction + Season + Depth, data = sampledf)
  ## Fraction R2= 0.11226; p-value = 0.001
  ## Season R2= 0.07723; p-value = 0.001
  ## Depth R2= 0.03532; p-value = 0.001
```



### PCoA of Whole Bacterial Community
```{r pcoa.bcc}
scaled.bray <- phyloseq::distance(physeq=scaled, method="bray")

scaled.pcoa <- ordinate(physeq = scaled, method = "PCoA", distance="bray")

### Color = Depth, Shape = Lake_Type
plot_ordination(
  physeq = scaled, 
  ordination = scaled.pcoa, 
  color = "Depth",
  shape = "Lake_Type",
  axes = c(1, 2),
  title = "Bray-Curtis Dissimilarity") +
  theme(axis.text.x = element_blank(), 
        axis.text.y = element_blank(),
        axis.title.x = element_text(face="bold", size=14),
        axis.title.y = element_text(face="bold", size=14),
        legend.title = element_text(size=12, face="bold"),
        legend.text = element_text(size=12),
        axis.ticks = element_blank(),
        plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm"), #top, right, bottom, left
        plot.title = element_text(size = 16, face="bold")) +
  guides(color = guide_legend(order=1, override.aes=list(size=3)),
         shape = guide_legend(order=2, override.aes=list(size=3)))


### Color = Lake_Type, Shape = Depth
bcc.pcoa.Lake_Type.depth <- plot_ordination(
  physeq = scaled, 
  ordination = scaled.pcoa, 
  color = "Lake_Type",
  shape = "Depth",
  axes = c(2, 3),
  title = "Bray-Curtis Dissimilarity") +
  scale_y_reverse()+
  scale_color_manual(values = c("indianred1","deepskyblue","mediumpurple2")) +
  theme(axis.text.x = element_blank(), 
        axis.text.y = element_blank(),
        axis.title.x = element_text(face="bold", size=14),
        axis.title.y = element_text(face="bold", size=14),
        legend.title = element_text(size=12, face="bold"),
        legend.text = element_text(size=12),
        axis.ticks = element_blank(),
        plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm"), #top, right, bottom, left
        plot.title = element_text(size = 16, face="bold")) +
  guides(color = guide_legend(order=1, override.aes=list(size=3)),
         shape = guide_legend(order=2, override.aes=list(size=3)))

### Facet = Lake_Type, Color = Season, Shape = Fraction
bcc.pcoa.facet <- plot_ordination(
  physeq = scaled, 
  ordination = scaled.pcoa, 
  color = "Season",
  shape = "Fraction",
  axes = c(1,2),
  title = "Bray-Curtis Dissimilarity") +
  facet_wrap(facets = ~Lake_Type) +
  scale_color_manual(values = c("indianred1","deepskyblue","mediumpurple2")) +
  theme(axis.text.x = element_blank(), 
        axis.text.y = element_blank(),
        axis.title.x = element_text(face="bold", size=14),
        axis.title.y = element_text(face="bold", size=14),
        legend.title = element_text(size=12, face="bold"),
        legend.text = element_text(size=12),
        axis.ticks = element_blank(),
        plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm"), #top, right, bottom, left
        plot.title = element_text(size = 16, face="bold")) +
  guides(color = guide_legend(order=1, override.aes=list(size=3)),
         shape = guide_legend(order=2, override.aes=list(size=3)))



```




### BCC Lake PCoA
```{r bcc.lake.pcoa}
### Laurentian
scaled.lau <- subset_samples(scaled, Lake_Type == "Laurentian")
scaled.lau.bray <- phyloseq::distance(physeq = scaled.lau, method = "bray")
scaled.lau.pcoa <- ordinate(physeq = scaled.lau, method = "PCoA", distance = "bray")


scaled.lau.ord <- plot_ordination(physeq = scaled.lau, ordination = scaled.lau.pcoa, axes = c(1, 2),
                title= "Bray-Curtis Dissimilarity") + 
  geom_point(size=2, aes(color = Season, shape = Depth,
                         fill = factor(ifelse(Fraction == "Free", NA, Season)),
                         alpha = Fraction)) +
  scale_shape_manual(values = c(21, 24, 22)) +
  scale_fill_discrete(na.value = "white", guide = "none") +
  scale_alpha_manual(values=c("Particle"=1, "Free"=1)) +
  #scale_y_reverse() + 
  theme(axis.text.x = element_blank(), 
        axis.text.y = element_blank(),
        axis.title.x = element_text(face="bold", size=10),
        axis.title.y = element_text(face="bold", size=10),
        legend.title = element_text(size=8, face="bold"),
        legend.text = element_text(size=8),
        axis.ticks = element_blank(),
        plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm"), #top, right, bottom, left
        plot.title = element_text(size = 10, face="bold")) +
  annotate("text", x= -0.28, y=0.45, label = "D", fontface = "bold", size=6, colour="black") +
  guides(colour=FALSE, shape=FALSE, alpha = FALSE) 
  #guides(color = guide_legend(order=1, title="Season", override.aes=list(size=3)),
  #       shape = guide_legend(order=2, title="Depth",override.aes=list(size=3)),
  #       alpha = guide_legend(order=3, title="Fraction",override.aes=list(size=3)))
# Particle = fill, Free = empty
# Note: for scale_fill_discrete to work, you MUST set scale_shape_manual to between 21-25
#        You also can't set colors manually



scaled.est <- subset_samples(scaled, Lake_Type == "Estuary")
scaled.est.bray <- phyloseq::distance(physeq = scaled.est, method = "bray")
scaled.est.pcoa <- ordinate(physeq = scaled.est, method = "PCoA", distance = "bray")
  
scaled.est.ord <- plot_ordination(physeq = scaled.est, ordination = scaled.est.pcoa, axes = c(1, 2),
                title= "Bray-Curtis Dissimilarity") + 
  geom_point(size=2, aes(color = Season, shape = Depth,
                         fill = factor(ifelse(Fraction == "Free", NA, Season)),
                         alpha = Fraction)) +
  scale_shape_manual(values = c(21, 24, 22)) +
  scale_fill_discrete(na.value = "white", guide = "none") +
  scale_alpha_manual(values=c("Particle"=1, "Free"=1, "Sediment"=0.5)) +
  theme(axis.text.x = element_blank(), 
        axis.text.y = element_blank(),
        axis.title.x = element_text(face="bold", size=10),
        axis.title.y = element_text(face="bold", size=10),
        legend.title = element_text(size=8, face="bold"),
        legend.text = element_text(size=8),
        axis.ticks = element_blank(),
        plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm"), #top, right, bottom, left
        plot.title = element_text(size = 10, face="bold")) +
  annotate("text", x= -0.24, y=0.45, label = "E", fontface = "bold", size=6, colour="black") +
  guides(colour=FALSE, shape=FALSE, alpha = FALSE) 
  #guides(color = guide_legend(order=1, title="Season", override.aes=list(size=3)),
  #       shape = guide_legend(order=2, title="Depth",override.aes=list(size=3)),
  #       alpha = guide_legend(order=3, title="Fraction",override.aes=list(size=3)))



  
  
scaled.inl <- subset_samples(scaled, Lake_Type == "Inland")
scaled.inl.bray <- phyloseq::distance(physeq = scaled.inl, method = "bray")
scaled.inl.pcoa <- ordinate(physeq = scaled.inl, method = "PCoA", distance = "bray")

scaled.inl.ord <- plot_ordination(physeq = scaled.inl, ordination = scaled.inl.pcoa, axes = c(1, 2),
                title= "Bray-Curtis Dissimilarity") + 
  geom_point(size=2, aes(color = Season, shape = Depth,
                         fill = factor(ifelse(Fraction == "Free", NA, Season)),
                         alpha = Fraction)) +
  scale_shape_manual(values = c(21, 24, 22)) +
  scale_fill_discrete(na.value = "white", guide = "none") +
  scale_alpha_manual(values=c("Particle"=1, "Free"=1)) +
  theme(axis.text.x = element_blank(), 
        axis.text.y = element_blank(),
        axis.title.x = element_text(face="bold", size=10),
        axis.title.y = element_text(face="bold", size=10),
        legend.title = element_text(size=8, face="bold"),
        legend.text = element_text(size=8),
        axis.ticks = element_blank(),
        plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm"), #top, right, bottom, left
        plot.title = element_text(size = 10, face="bold")) +
  annotate("text", x= -0.37, y=0.6, label = "F", fontface = "bold", size=6, colour="black") +
  guides(colour=FALSE, shape=FALSE, alpha = FALSE) 
  #guides(color = guide_legend(order=1, title="Season", override.aes=list(size=3)),
  #       shape = guide_legend(order=2, title="Depth",override.aes=list(size=3)),
  #       alpha = guide_legend(order=3, title="Fraction",override.aes=list(size=3)))




# tiff("bcc.pcoa.tiff", width = 15, height = 10, units = "cm", res = 500)
# grid.newpage()
# pushViewport(viewport(layout=(grid.layout(1,3, widths=c(1,1,1), heights=c(1,1,1)))))
# print(scaled.lau.ord, vp=viewport(layout.pos.row=1, layout.pos.col=1))
# print(scaled.est.ord, vp=viewport(layout.pos.row=1, layout.pos.col=2))
# print(scaled.inl.ord, vp=viewport(layout.pos.row=1, layout.pos.col=3))
# dev.off()
```


### MAKE BIG PCOA FIGURE
```{r big.pcoa}
tiff("pcoa.tiff", width = 17.8, height = 17.8, units = "cm", res = 600)
grid.newpage()
pushViewport(viewport(layout=(grid.layout(3,5, widths=c(1.5,19,19,19,7.5), heights=c(0.7,14,14)))))
grid.text("Laurentian", just=c(0.35, 0.5), vp=viewport(layout.pos.row=1, layout.pos.col=2), gp=gpar(fontsize=12,fontface="bold"))
grid.text("Estuary", just=c(0.35, 0.5), vp=viewport(layout.pos.row=1, layout.pos.col=3), gp=gpar(fontsize=12,fontface="bold"))
grid.text("Inland",  just=c(0.35, 0.5), vp=viewport(layout.pos.row=1, layout.pos.col=4), gp=gpar(fontsize=12,fontface="bold"))
grid.text("Verrucomicrobial Community", rot=90, vp=viewport(layout.pos.row=2, layout.pos.col=1), gp=gpar(fontsize=12,fontface="bold"))
grid.text("Whole Bacterial Community", rot=90, vp=viewport(layout.pos.row=3, layout.pos.col=1), gp=gpar(fontsize=12,fontface="bold"))
print(Ver.lau.ord, vp=viewport(layout.pos.row=2, layout.pos.col=2))
print(Ver.est.ord, vp=viewport(layout.pos.row=2, layout.pos.col=3))
print(Ver.inl.ord, vp=viewport(layout.pos.row=2, layout.pos.col=4))
print(scaled.lau.ord, vp=viewport(layout.pos.row=3, layout.pos.col=2))
print(scaled.est.ord, vp=viewport(layout.pos.row=3, layout.pos.col=3))
print(scaled.inl.ord, vp=viewport(layout.pos.row=3, layout.pos.col=4))
legend$vp <- viewport(layout.pos.row = 1:3, layout.pos.col = 5)
grid.draw(legend)
dev.off()
```


### Compare Whole Bacterial Community with Verruco Community
```{r bcc.ver.pcoa}
tiff("bcc.ver.compare2.tiff", width = 14, height = 10, units = "in", res = 500)
grid.newpage()
pushViewport(viewport(layout=(grid.layout(2,2, widths=c(1,1), heights=c(1.5,1.5)))))
print(bray.pcoa.Lake_Type.depth, vp=viewport(layout.pos.row=1, layout.pos.col=1))
print(bray.pcoa.facet, vp=viewport(layout.pos.row=1, layout.pos.col=2))
print(bcc.pcoa.Lake_Type.depth, vp=viewport(layout.pos.row=2, layout.pos.col=1))
print(bcc.pcoa.facet, vp=viewport(layout.pos.row=2, layout.pos.col=2))
dev.off()

```

### PERMANOVA for Whole Bacterial Community
```{r adonis.bcc}
bcc.sampledf <- data.frame(sample_data(scaled))

##### Depth #####
adonis(scaled.bray ~ Depth, data=bcc.sampledf)
  ## R2 = 0.09531
  ## p-value = 0.001
beta.tax = betadisper(d=scaled.bray, group=bcc.sampledf$Depth)
p <- permutest(beta.tax)
p$tab
  # p-value = 0.001
  # F = 18.0238 

# Top vs. Bottom
adonis.scaled.tb <- subset_samples(physeq = scaled, Depth != "DCM")
adonis.scaled.tb <- subset_samples(physeq = adonis.scaled.tb, Depth != "Sediment")
adonis.scaled.tb.bray <- phyloseq::distance(physeq = adonis.scaled.tb, method = "bray")
adonis.scaled.tb <- data.frame(sample_data(adonis.scaled.tb))
adonis(adonis.scaled.tb.bray ~ Depth, data=adonis.scaled.tb)
  # p-value = 0.002
  # R2 = 0.02071
beta.tax = betadisper(d=adonis.scaled.tb.bray, group=adonis.scaled.tb$Depth)
p <- permutest(beta.tax)
p$tab
  # p-value = 0.003
  # F = 9.793284

# Top vs. Sediment
adonis.scaled.ts <- subset_samples(physeq = scaled, Depth != "DCM")
adonis.scaled.ts <- subset_samples(physeq = adonis.scaled.ts, Depth != "Bottom")
adonis.scaled.ts.bray <- phyloseq::distance(physeq = adonis.scaled.ts, method = "bray")
adonis.scaled.ts <- data.frame(sample_data(adonis.scaled.ts))
adonis(adonis.scaled.ts.bray ~ Depth, data=adonis.scaled.ts)
  # p-value = 0.001
  # R2 = 0.14538
beta.tax = betadisper(d=adonis.scaled.ts.bray, group=adonis.scaled.ts$Depth)
p <- permutest(beta.tax)
p$tab
  # p-value = 0.001
  # F = 19.84845 

# Bottom vs. Sediment
adonis.scaled.bs <- subset_samples(physeq = scaled, Depth != "DCM")
adonis.scaled.bs <- subset_samples(physeq = adonis.scaled.bs, Depth != "Top")
adonis.scaled.bs.bray <- phyloseq::distance(physeq = adonis.scaled.bs, method = "bray")
adonis.scaled.bs <- data.frame(sample_data(adonis.scaled.bs))
adonis(adonis.scaled.bs.bray ~ Depth, data=adonis.scaled.bs)
  # p-value = 0.001
  # R2 = 0.12809
beta.tax = betadisper(d=adonis.scaled.bs.bray, group=adonis.scaled.bs$Depth)
p <- permutest(beta.tax)
p$tab
  # p-value = 0.001
  # F = 31.08996 


##### Lake_Type #####
adonis(scaled.bray ~ Lake_Type, data=bcc.sampledf)
  ## R2 = 0.09902
  ## p-value = 0.001
beta.tax = betadisper(scaled.bray, bcc.sampledf$Lake_Type)
p <- permutest(beta.tax)
p$tab
  ## p-value = 0.003
  ## F = 5.292305

# Laurentian vs. Estuary
adonis.scaled.le <- subset_samples(physeq = scaled, Lake_Type != "Inland")
adonis.scaled.le.bray <- phyloseq::distance(physeq = adonis.scaled.le, method = "bray")
adonis.scaled.le <- data.frame(sample_data(adonis.scaled.le))
adonis(adonis.scaled.le.bray ~ Lake_Type, data=adonis.scaled.le)
  # p-value = 0.001
  # R2 = 0.09594
beta.tax = betadisper(d=adonis.scaled.le.bray, group=adonis.scaled.le$Lake_Type)
p <- permutest(beta.tax)
p$tab
  # p-value = 0.082
  # F = 3.061581

# Laurentian vs. Inland
adonis.scaled.li <- subset_samples(physeq = Ver.rel, Lake_Type != "Estuary")
adonis.scaled.li.bray <- phyloseq::distance(physeq = adonis.scaled.li, method = "bray")
adonis.scaled.li <- data.frame(sample_data(adonis.scaled.li))
adonis(adonis.scaled.li.bray ~ Lake_Type, data=adonis.scaled.li)
  # p-value = 0.001
  # R2 = 0.08015
beta.tax = betadisper(d=adonis.scaled.li.bray, group=adonis.scaled.li$Lake_Type)
p <- permutest(beta.tax)
p$tab
  # p-value = 0.481
  # F = 0.4470369

# Estuary vs. Inland
adonis.ei <- subset_samples(physeq = Ver.rel, Lake_Type != "Laurentian")
adonis.ei.bray <- phyloseq::distance(physeq = adonis.ei, method = "bray")
adonis.ei <- data.frame(sample_data(adonis.ei))
adonis(adonis.ei.bray ~ Lake_Type, data=adonis.ei)
  # p-value = 0.001
  # R2 = 0.07186
beta.tax = betadisper(d=adonis.ei.bray, group=adonis.ei$Lake_Type)
p <- permutest(beta.tax)
p$tab
  # p-value = 0.763
  # F = 0.09847275 




##### Fraction #####
adonis(scaled.bray ~ Fraction, data=bcc.sampledf)
  ## p-value = 0.001
  ## R2 = 0.21101
beta.tax = betadisper(scaled.bray, bcc.sampledf$Fraction)
p <- permutest(beta.tax)
p$tab
  ## P-value = 0.001
  ## F = 60.67025

# Particle vs. Free
adonis.scaled.pf <- subset_samples(physeq = scaled, Fraction != "Sediment")
adonis.scaled.pf.bray <- phyloseq::distance(physeq = adonis.scaled.pf, method = "bray")
adonis.scaled.pf <- data.frame(sample_data(adonis.scaled.pf))
adonis(adonis.scaled.pf.bray ~ Fraction, data=adonis.scaled.pf)
  # p-value = 0.001
  # R2 = 0.14928
beta.tax = betadisper(d=adonis.scaled.pf.bray, group=adonis.scaled.pf$Fraction)
p <- permutest(beta.tax)
p$tab
  # p-value = 0.001
  # F = 105.9411

# Particle vs. Sediment
adonis.scaled.ps <- subset_samples(physeq = scaled, Fraction != "Free")
adonis.scaled.ps.bray <- phyloseq::distance(physeq = adonis.scaled.ps, method = "bray")
adonis.scaled.ps <- data.frame(sample_data(adonis.scaled.ps))
adonis(adonis.scaled.ps.bray ~ Fraction, data=adonis.scaled.ps)
  # p-value = 0.001
  # R2 = 0.12068
beta.tax = betadisper(d=adonis.scaled.ps.bray, group=adonis.scaled.ps$Fraction)
p <- permutest(beta.tax)
p$tab
  # p-value = 0.001
  # F = 45.63044

# Free vs. Sediment
adonis.scaled.fs <- subset_samples(physeq = scaled, Fraction != "Particle")
adonis.scaled.fs.bray <- phyloseq::distance(physeq = adonis.scaled.fs, method = "bray")
adonis.scaled.fs <- data.frame(sample_data(adonis.scaled.fs))
adonis(adonis.scaled.fs.bray ~ Fraction, data=adonis.scaled.fs)
  # p-value = 0.001
  # R2 = 0.22202
beta.tax = betadisper(d=adonis.scaled.fs.bray, group=adonis.scaled.fs$Fraction)
p <- permutest(beta.tax)
p$tab
  # p-value = 0.106
  # F = 2.668599


##### Season #####
adonis(scaled.bray ~ Season, data=bcc.sampledf)
  ## p-value = 0.001
  ## R2 = 0.0832
beta.tax = betadisper(scaled.bray, bcc.sampledf$Season)
p <- permutest(beta.tax)
p$tab
  ## P-value = 0.025
  ## F = 4.260132

# Spring vs. Summer
adonis.scaled.ss <- subset_samples(physeq = scaled, Season != "Fall")
adonis.scaled.ss.bray <- phyloseq::distance(physeq = adonis.scaled.ss, method = "bray")
adonis.scaled.ss <- data.frame(sample_data(adonis.scaled.ss))
adonis(adonis.scaled.ss.bray ~ Season, data=adonis.scaled.ss)
  # p-value = 0.001
  # R2 = 0.07992
beta.tax = betadisper(d=adonis.scaled.ss.bray, group=adonis.scaled.ss$Fraction)
p <- permutest(beta.tax)
p$tab
  # p-value = 0.001
  # F = 50.16244

# Spring vs. Fall
adonis.scaled.sf <- subset_samples(physeq = scaled, Season != "Summer")
adonis.scaled.sf.bray <- phyloseq::distance(physeq = adonis.scaled.sf, method = "bray")
adonis.scaled.sf <- data.frame(sample_data(adonis.scaled.sf))
adonis(adonis.scaled.sf.bray ~ Season, data=adonis.scaled.sf)
  # p-value = 0.001
  # R2 = 0.10023 
beta.tax = betadisper(d=adonis.scaled.sf.bray, group=adonis.scaled.sf$Fraction)
p <- permutest(beta.tax)
p$tab
  # p-value = 0.001
  # F = 30.20449 

# Summer vs. Fall
adonis.scaled.suf <- subset_samples(physeq = scaled, Season != "Spring")
adonis.scaled.suf.bray <- phyloseq::distance(physeq = adonis.scaled.suf, method = "bray")
adonis.scaled.suf <- data.frame(sample_data(adonis.scaled.suf))
adonis(adonis.scaled.suf.bray ~ Season, data=adonis.scaled.suf)
  # p-value = 0.001
  # R2 = 0.02862
beta.tax = betadisper(d=adonis.scaled.suf.bray, group=adonis.scaled.suf$Fraction)
p <- permutest(beta.tax)
p$tab
  # p-value = 0.001
  # F = 32.70444


##### Nested adonis #####
adonis(scaled.bray ~ Lake_Type + Depth + Fraction + Season, data = bcc.sampledf)
  ## Lake Type R2 = 0.09902
  ## Depth R2 = 0.09387
  ## Fraction R2 = 0.13366
  ## Season = 0.06705
  ## All p-values = 0.001

# if you rearrange the variables (and include sed as part of frac)....
  ## Fraction R2 = 0.21101
  ## Lake_Type R2 = 0.09730
  ## Season R2 = 0.06683
  ## Depth R2 = 0.01848
  ## All p-values = 0.001

```




### BCC Lake PERMANOVA
```{r bcc.lake.permanova}
########## LAURENTIAN ##########
sampledf <- data.frame(sample_data(scaled.lau))

### Depth
adonis(scaled.lau.bray ~ Depth, data=sampledf)
  # R2 = 0.03723
  # p-value = 0.238
beta.tax = betadisper(d=scaled.lau.bray, group=sampledf$Depth)
p <- permutest(beta.tax)
p$tab
  # p-value = 0.849
  # F = 0.03704203

### Season
adonis(scaled.lau.bray ~ Season, data=sampledf)
  # R2 = 0.23951
  # p-value = 0.001
beta.tax = betadisper(d=scaled.lau.bray, group=sampledf$Season)
p <- permutest(beta.tax)
p$tab
  # p-value = 0.141
  # F = 2.141275

# Spring vs. Summer
adonis.ss <- subset_samples(physeq = scaled.lau, Season != "Fall")
adonis.ss.bray <- phyloseq::distance(physeq = adonis.ss, method = "bray")
adonis.ss <- data.frame(sample_data(adonis.ss))
adonis(adonis.ss.bray ~ Season, data=adonis.ss)
  # R2 = 0.2255
  # p-value = 0.001
beta.tax = betadisper(d=adonis.ss.bray, group=adonis.ss$Season)
p <- permutest(beta.tax)
p$tab
  # p-value = 0.121
  # F = 2.479235

# Spring vs. Fall
adonis.spf <- subset_samples(physeq = scaled.lau, Season != "Summer")
adonis.spf.bray <- phyloseq::distance(physeq = adonis.spf, method = "bray")
adonis.spf <- data.frame(sample_data(adonis.spf))
adonis(adonis.spf.bray ~ Season, data=adonis.spf)
  # R2 = 0.23554
  # p-value = 0.001
beta.tax = betadisper(d=adonis.spf.bray, group=adonis.spf$Season)
p <- permutest(beta.tax)
p$tab
  # p-value = 0.073
  # F = 3.621429

# Summer vs. Fall
adonis.suf <- subset_samples(physeq = scaled.lau, Season != "Spring")
adonis.suf.bray <- phyloseq::distance(physeq = adonis.suf, method = "bray")
adonis.suf <- data.frame(sample_data(adonis.suf))
adonis(adonis.suf.bray ~ Season, data=adonis.suf)
  # R2 = 0.11285
  # p-value = 0.047
beta.tax = betadisper(d=adonis.suf.bray, group=adonis.suf$Season)
p <- permutest(beta.tax)
p$tab
  # p-value = 0.918
  # F = 0.01387363

### Fraction
adonis(scaled.lau.bray ~ Fraction, data=sampledf)
  # R2 = 0.34359
  # p-value = 0.001
beta.tax = betadisper(d=scaled.lau.bray, group=sampledf$Fraction)
p <- permutest(beta.tax)
p$tab
  # p-value = 0.001
  # F = 52.05048

##### Nested adonis #####
adonis(scaled.lau.bray ~ Fraction + Season + Depth, data = sampledf)
  ## Fraction R2= 0.34359; p-value = 0.001
  ## Season R2= 0.24164; p-value = 0.001
  ## Depth R2= 0.03411; p-value = 0.025


########## ESTUARY ##########
sampledf <- data.frame(sample_data(scaled.est))

### Depth
adonis(scaled.est.bray ~ Depth, data=sampledf)
  # R2 = 0.30522
  # p-value = 0.001
beta.tax = betadisper(d=scaled.est.bray, group=sampledf$Depth)
p <- permutest(beta.tax)
p$tab
  # p-value = 0.207
  # F = 1.539821

# Water vs. Sediment
adonis(scaled.est.bray ~ Source, data=sampledf)
  # R2 = 0.28846
  # p-value = 0.001
beta.tax = betadisper(d=scaled.est.bray, group=sampledf$Source)
p <- permutest(beta.tax)
p$tab
  # p-value = 0.084
  # F = 3.271663

# Top vs. Bottom
adonis.tb <- subset_samples(physeq = scaled.est, Depth != "Sediment")
adonis.tb.bray <- phyloseq::distance(physeq = adonis.tb, method = "bray")
adonis.tb <- data.frame(sample_data(adonis.tb))
adonis(adonis.tb.bray ~ Depth, data=adonis.tb)
  # R2 = 0.02693
  # p-value = 0.189
beta.tax = betadisper(d=adonis.tb.bray, group=adonis.tb$Depth)
p <- permutest(beta.tax)
p$tab
  # p-value = 0.501
  # F = 0.4961036

# Top vs. Sediment
adonis.ts <- subset_samples(physeq = scaled.est, Depth != "Bottom")
adonis.ts.bray <- phyloseq::distance(physeq = adonis.ts, method = "bray")
adonis.ts <- data.frame(sample_data(adonis.ts))
adonis(adonis.ts.bray ~ Depth, data=adonis.ts)
  # R2 = 0.40019
  # p-value = 0.001
beta.tax = betadisper(d=adonis.ts.bray, group=adonis.ts$Depth)
p <- permutest(beta.tax)
p$tab
  # p-value = 0.245
  # F = 1.410866

# Bottom vs. Sediment
adonis.bs <- subset_samples(physeq = scaled.est, Depth != "Top")
adonis.bs.bray <- phyloseq::distance(physeq = adonis.bs, method = "bray")
adonis.bs <- data.frame(sample_data(adonis.bs))
adonis(adonis.bs.bray ~ Depth, data=adonis.bs)
  # R2 = 0.37597
  # p-value = 0.001
beta.tax = betadisper(d=adonis.bs.bray, group=adonis.bs$Depth)
p <- permutest(beta.tax)
p$tab
  # p-value = 0.06
  # F = 3.567448

### Season
adonis(scaled.est.bray ~ Season, data=sampledf)
  # R2 = 0.17415
  # p-value = 0.001
beta.tax = betadisper(d=scaled.est.bray, group=sampledf$Season)
p <- permutest(beta.tax)
p$tab
  # p-value = 0.756
  # F = 0.2654727

# Spring vs. Summer
adonis.ss <- subset_samples(physeq = scaled.est, Season != "Fall")
adonis.ss.bray <- phyloseq::distance(physeq = adonis.ss, method = "bray")
adonis.ss <- data.frame(sample_data(adonis.ss))
adonis(adonis.ss.bray ~ Season, data=adonis.ss)
  # R2 = 0.18032
  # p-value = 0.001
beta.tax = betadisper(d=adonis.ss.bray, group=adonis.ss$Season)
p <- permutest(beta.tax)
p$tab
  # p-value = 0.856
  # F = 0.02931041

# Spring vs. Fall
adonis.spf <- subset_samples(physeq = scaled.est, Season != "Summer")
adonis.spf.bray <- phyloseq::distance(physeq = adonis.spf, method = "bray")
adonis.spf <- data.frame(sample_data(adonis.spf))
adonis(adonis.spf.bray ~ Season, data=adonis.spf)
  # R2 = 0.18804
  # p-value = 0.001
beta.tax = betadisper(d=adonis.spf.bray, group=adonis.spf$Season)
p <- permutest(beta.tax)
p$tab
  # p-value = 0.521
  # F = 0.4368101

# Summer vs. Fall
adonis.suf <- subset_samples(physeq = scaled.est, Season != "Spring")
adonis.suf.bray <- phyloseq::distance(physeq = adonis.suf, method = "bray")
adonis.suf <- data.frame(sample_data(adonis.suf))
adonis(adonis.suf.bray ~ Season, data=adonis.suf)
  # R2 = 0.05209
  # p-value = 0.042
beta.tax = betadisper(d=adonis.suf.bray, group=adonis.suf$Season)
p <- permutest(beta.tax)
p$tab
  # p-value = 0.594
  # F = 0.2960811

### Fraction
adonis(scaled.est.bray ~ Fraction, data=sampledf)
  # R2 = 0.41303
  # p-value = 0.001
beta.tax = betadisper(d=scaled.est.bray, group=sampledf$Fraction)
p <- permutest(beta.tax)
p$tab
  # p-value = 0.045
  # F = 3.458586

# Particle vs. Free
adonis.frac <- subset_samples(physeq = scaled.est, Fraction != "Sediment")
adonis.frac.bray <- phyloseq::distance(physeq = adonis.frac, method = "bray")
adonis.frac <- data.frame(sample_data(adonis.frac))
adonis(adonis.frac.bray ~ Fraction, data=adonis.frac)
  # R2 = 0.20022
  # p-value = 0.001
beta.tax = betadisper(d=adonis.frac.bray, group=adonis.frac$Fraction)
p <- permutest(beta.tax)
p$tab
  # p-value = 0.009
  # F = 6.297787

# Particle vs. Sediment
adonis.ps <- subset_samples(physeq = scaled.est, Fraction != "Free")
adonis.ps.bray <- phyloseq::distance(physeq = adonis.ps, method = "bray")
adonis.ps <- data.frame(sample_data(adonis.ps))
adonis(adonis.ps.bray ~ Fraction, data=adonis.ps)
  # R2 = 0.37385
  # p-value = 0.001
beta.tax = betadisper(d=adonis.ps.bray, group=adonis.ps$Fraction)
p <- permutest(beta.tax)
p$tab
  # p-value = 0.212
  # F = 1.47037

# Free vs. Sediment
adonis.fs <- subset_samples(physeq = scaled.est, Fraction != "Particle")
adonis.fs.bray <- phyloseq::distance(physeq = adonis.fs, method = "bray")
adonis.fs <- data.frame(sample_data(adonis.fs))
adonis(adonis.fs.bray ~ Fraction, data=adonis.fs)
  # R2 = 0.51092
  # p-value = 0.001
beta.tax = betadisper(d=adonis.fs.bray, group=adonis.fs$Fraction)
p <- permutest(beta.tax)
p$tab
  # p-value = 0.43
  # F = 0.5958934


##### Nested adonis #####
adonis(scaled.est.bray ~ Depth + Season + Fraction, data = sampledf)
  ## Depth R2= 0.30522; p-value = 0.001
  ## Season R2= 0.17341; p-value = 0.001
  ## Fraction R2= 0.12457; p-value = 0.001




########## INLAND ##########
sampledf <- data.frame(sample_data(scaled.inl))

### Depth
adonis(scaled.inl.bray ~ Depth, data=sampledf)
  # R2 = 0.03975
  # p-value = 0.001
beta.tax = betadisper(d=scaled.inl.bray, group=sampledf$Depth)
p <- permutest(beta.tax)
p$tab
  # p-value = 0.001
  # F = 25.05928

### Season
adonis(scaled.inl.bray ~ Season, data=sampledf)
  # R2 = 0.09735
  # p-value = 0.001
beta.tax = betadisper(d=scaled.inl.bray, group=sampledf$Season)
p <- permutest(beta.tax)
p$tab
  # p-value = 0.001
  # F = 15.04692

# Spring vs. Summer
adonis.ss <- subset_samples(physeq = scaled.inl, Season != "Fall")
adonis.ss.bray <- phyloseq::distance(physeq = adonis.ss, method = "bray")
adonis.ss <- data.frame(sample_data(adonis.ss))
adonis(adonis.ss.bray ~ Season, data=adonis.ss)
  # R2 = 0.08529
  # p-value = 0.001
beta.tax = betadisper(d=adonis.ss.bray, group=adonis.ss$Season)
p <- permutest(beta.tax)
p$tab
  # p-value = 0.001
  # F = 30.3471

# Spring vs. Fall
adonis.spf <- subset_samples(physeq = scaled.inl, Season != "Summer")
adonis.spf.bray <- phyloseq::distance(physeq = adonis.spf, method = "bray")
adonis.spf <- data.frame(sample_data(adonis.spf))
adonis(adonis.spf.bray ~ Season, data=adonis.spf)
  # R2 = 0.20982
  # p-value = 0.001
beta.tax = betadisper(d=adonis.spf.bray, group=adonis.spf$Season)
p <- permutest(beta.tax)
p$tab
  # p-value = 0.029
  # F = 5.219433

# Summer vs. Fall
adonis.suf <- subset_samples(physeq = scaled.inl, Season != "Spring")
adonis.suf.bray <- phyloseq::distance(physeq = adonis.suf, method = "bray")
adonis.suf <- data.frame(sample_data(adonis.suf))
adonis(adonis.suf.bray ~ Season, data=adonis.suf)
  # R2 = 0.02688
  # p-value = 0.002
beta.tax = betadisper(d=adonis.suf.bray, group=adonis.suf$Season)
p <- permutest(beta.tax)
p$tab
  # p-value = 0.065
  # F = 4.040812

### Fraction
adonis(scaled.inl.bray ~ Fraction, data=sampledf)
  # R2 = 0.17221
  # p-value = 0.001
beta.tax = betadisper(d=scaled.inl.bray, group=sampledf$Fraction)
p <- permutest(beta.tax)
p$tab
  # p-value = 0.001
  # F = 26.48775

##### Nested adonis #####
adonis(scaled.inl.bray ~ Fraction + Season + Depth, data = sampledf)
  ## Fraction R2= 0.17221; p-value = 0.001
  ## Season R2= 0.09735; p-value = 0.001
  ## Depth R2= 0.04004; p-value = 0.001
```




### Procrustes Analyses (Verruco vs. Whole Bacterial community)
```{r procrustes}
procrust <- procrustes(scaled.pcoa$vectors[,1:2], Ver.pcoa$vectors[,1:2], scale=FALSE,
                       symmetric=FALSE)
  # Calculates procrustes based upon coordinates on axes 1 & 2 of the Bray-Curtis PCoA
  # Sum of squares = 12.43
plot(procrust)
plot(procrust,kind=2)
pro <- protest(scaled.pcoa$vectors[,1:2], Ver.pcoa$vectors[,1:2])
  # p-value = 0.001
  # m12^2 (sum of squares) = 0.3896
  # correlation = 0.7813
plot(pro)


procrust2 <- procrustes(scaled.pcoa$vectors[,1:3], Ver.pcoa$vectors[,1:3], scale=FALSE,
                       symmetric=FALSE)
  # Calculates procrustes based upon coordinates on axes 1 & 2 of the Bray-Curtis PCoA
  # Sum of squares = 11.72
plot(procrust2)
plot(procrust2,kind=2)
pro2 <- protest(scaled.pcoa$vectors[,1:3], Ver.pcoa$vectors[,1:3])
  # p-value = 0.001
  # m12^2 (sum of squares) = 0.2891
  # correlation = 0.8431
plot(pro2)


procrust3 <- procrustes(scaled.pcoa$vectors[,2:3], Ver.pcoa$vectors[,2:3], scale=FALSE,
                       symmetric=FALSE)
  # Calculates procrustes based upon coordinates on axes 2 & 3 of the Bray-Curtis PCoA
  # Sum of squares = 7.538
plot(procrust3)
plot(procrust3,kind=2)
pro3 <- protest(scaled.pcoa$vectors[,2:3], Ver.pcoa$vectors[,2:3])
  # p-value = 0.001
  # m12^2 (sum of squares) = 0.3058
  # correlation = 0.8332
plot(pro2)




### Prepare to make a prettier Procrustes plot
# http://stackoverflow.com/questions/30325739/ggplot2-for-procrustes-rotation-in-vegan
pro.df <- data.frame(Dimension1=procrust2$Yrot[,1], Dimension2=procrust2$Yrot[,2],
                     xDimension1=procrust2$X[,1], xDimension2=procrust2$X[,2])
pro.df$Depth <- sample_data(Ver)$Depth
pro.df$Lake_Type <- sample_data(Ver)$Lake_Type
pro.df$Fraction <- sample_data(Ver)$Fraction
pro.df$Season <- sample_data(Ver)$Season

### Make a prettier Procrustes plot
# Basic
#tiff("procrustes.tiff", width = 8, height = 5, units = "in", res = 500)
ggplot(pro.df) +
  geom_point(aes(x=Dimension1, y=Dimension2)) +
  geom_point(aes(x=xDimension1, y=xDimension2)) +
  geom_segment(aes(x=Dimension2, y=Dimension2, xend=xDimension1, yend=xDimension2),
               arrow=arrow(length=unit(0.2,"cm")), color="burlywood", alpha=0.5)
#dev.off()

### Facet = Lake_Type, Color = Depth, Shape = Fraction
#tiff("procrustes.facet1.tiff", width = 8, height = 5, units = "in", res = 500)
ggplot(pro.df) +
  geom_point(aes(x=Dimension1, y=Dimension2, color=Depth, shape=Fraction)) +
  geom_point(aes(x=xDimension1, y=xDimension2, color=Depth, shape=Fraction)) +
  geom_segment(aes(x=Dimension2, y=Dimension2, xend=xDimension1, yend=xDimension2),
               arrow=arrow(length=unit(0.2,"cm")), color="burlywood", alpha=0.5) +
  facet_grid(~Lake_Type) +
  scale_color_manual(values = c("indianred1","green3","mediumpurple2", "skyblue")) +
  theme(axis.text.x = element_blank(), 
        axis.text.y = element_blank(),
        axis.title.x = element_text(face="bold", size=14),
        axis.title.y = element_text(face="bold", size=14),
        legend.title = element_text(size=12, face="bold"),
        legend.text = element_text(size=12),
        axis.ticks = element_blank(),
        plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm"), #top, right, bottom, left
        plot.title = element_text(size = 16, face="bold")) +
  guides(color = guide_legend(order=1, override.aes=list(size=3)),
         shape = guide_legend(order=2, override.aes=list(size=3)))
#dev.off()

### Facet = Lake_Type, Color = Fraction, Shape = Season
#tiff("procrustes.facet2.tiff", width = 8, height = 5, units = "in", res = 500)
ggplot(pro.df) +
  geom_point(aes(x=Dimension1, y=Dimension2, color=Fraction, shape=Season)) +
  geom_point(aes(x=xDimension1, y=xDimension2, color=Fraction, shape=Season)) +
  geom_segment(aes(x=Dimension2, y=Dimension2, xend=xDimension1, yend=xDimension2),
               arrow=arrow(length=unit(0.2,"cm")), color="burlywood", alpha=0.5) +
  facet_grid(~Lake_Type) +
  scale_color_manual(values = c("indianred1","skyblue","mediumpurple2")) +
  theme(axis.text.x = element_blank(), 
        axis.text.y = element_blank(),
        axis.title.x = element_text(face="bold", size=14),
        axis.title.y = element_text(face="bold", size=14),
        legend.title = element_text(size=12, face="bold"),
        legend.text = element_text(size=12),
        axis.ticks = element_blank(),
        plot.margin = unit(c(0.1, 0.1, 0.1, 0.1), "cm"), #top, right, bottom, left
        plot.title = element_text(size = 16, face="bold")) +
  guides(color = guide_legend(order=1, override.aes=list(size=3)),
         shape = guide_legend(order=2, override.aes=list(size=3)))
#dev.off()




```













### Lake Procrustes Analyses (Verruco vs. Whole Bacterial community)
```{r lake.procrustes}

########## LAURENTIAN ##########
procrust <- procrustes(scaled.lau.pcoa$vectors[,1:2], Ver.lau.pcoa$vectors[,1:2], scale=FALSE,
                       symmetric=FALSE)
  # Calculates procrustes based upon coordinates on axes 1 & 2 of the Bray-Curtis PCoA
  # Sum of squares = 2.446
    # Sum of squares = 1.901
plot(procrust)
plot(procrust,kind=2)
pro <- protest(scaled.lau.pcoa$vectors[,1:2], Ver.lau.pcoa$vectors[,1:2])
  # p-value = 0.001
  # m12^2 (sum of squares) = 0.4662
    # m12^2 (sum of squares) = 0.3399
  # correlation = 0.7306
    # correlation = 0.8125
plot(pro)


procrust2 <- procrustes(scaled.lau.pcoa$vectors[,1:3], Ver.lau.pcoa$vectors[,1:3], scale=FALSE,
                       symmetric=FALSE)
  # Calculates procrustes based upon coordinates on axes 1 & 2 of the Bray-Curtis PCoA
  # Sum of squares = 2.08
    # sum of squares = 2.34
plot(procrust2)
plot(procrust2,kind=2)
pro2 <- protest(scaled.lau.pcoa$vectors[,1:3], Ver.lau.pcoa$vectors[,1:3])
  # p-value = 0.001
  # m12^2 (sum of squares) = 0.33
    # m12^2 (sum of squares) = 0.3417
  # correlation = 0.8185
    # correlation = 0.8114
plot(pro2)







########## ESTUARY ##########
procrust <- procrustes(scaled.est.pcoa$vectors[,1:2], Ver.est.pcoa$vectors[,1:2], scale=FALSE,
                       symmetric=FALSE)
  # Calculates procrustes based upon coordinates on axes 1 & 2 of the Bray-Curtis PCoA
  # Sum of squares = 0.9803
    # sum of squares = 1.05
plot(procrust)
plot(procrust,kind=2)
pro <- protest(scaled.est.pcoa$vectors[,1:2], Ver.est.pcoa$vectors[,1:2])
  # p-value = 0.001
  # m12^2 (sum of squares) = 0.08122
    # m12^2 (sum of squares) = 0.09223
  # correlation = 0.9585
    # correlation = 0.9528
plot(pro)


procrust2 <- procrustes(scaled.est.pcoa$vectors[,1:3], Ver.est.pcoa$vectors[,1:3], scale=FALSE,
                       symmetric=FALSE)
  # Calculates procrustes based upon coordinates on axes 1 & 2 of the Bray-Curtis PCoA
  # Sum of squares = 1.315
    # sum of squares = 1.257
plot(procrust2)
plot(procrust2,kind=2)
pro2 <- protest(scaled.est.pcoa$vectors[,1:3], Ver.est.pcoa$vectors[,1:3])
  # p-value = 0.001
  # m12^2 (sum of squares) = 0.09107
    # m12^2 (sum of squares) = 0.08774
  # correlation = 0.9534
    # correlation = 0.9551
plot(pro2)






########## INLAND ##########
procrust <- procrustes(scaled.inl.pcoa$vectors[,1:2], Ver.inl.pcoa$vectors[,1:2], scale=FALSE,
                       symmetric=FALSE)
  # Calculates procrustes based upon coordinates on axes 1 & 2 of the Bray-Curtis PCoA
  # Sum of squares = 5.582
    # sum of squares = 5.454
plot(procrust)
plot(procrust,kind=2)
pro <- protest(scaled.inl.pcoa$vectors[,1:2], Ver.inl.pcoa$vectors[,1:2])
  # p-value = 0.001
  # m12^2 (sum of squares) = 0.3757
    # m12^2 (sum of squares) = 0.3649
  # correlation = 0.7901
    # correlation = 0.7969
plot(pro)


procrust2 <- procrustes(scaled.inl.pcoa$vectors[,1:3], Ver.inl.pcoa$vectors[,1:3], scale=FALSE,
                       symmetric=FALSE)
  # Calculates procrustes based upon coordinates on axes 1 & 2 of the Bray-Curtis PCoA
  # Sum of squares = 7.864
    # sum of squares = 8.086
plot(procrust2)
plot(procrust2,kind=2)
pro2 <- protest(scaled.inl.pcoa$vectors[,1:3], Ver.inl.pcoa$vectors[,1:3])
  # p-value = 0.001
  # m12^2 (sum of squares) = 0.4154
    # m12^2 (sum of squares) = 0.4168
  # correlation = 0.7646
    # correlation = 0.7637
plot(pro2)




```




### RELATIVE ABUNDANCE UNSTACKED BAR PLOT
```{r}
classVer <- tax_glom(Ver, taxrank = "Class")

# Calculate relative abundance
classVer.rel = transform_sample_counts(classVer, function(x){x/sum(x)})

# If a phylum has <1% rel abun, change that to 0%
otu_table(classVer.rel)[otu_table(classVer.rel)<.01] <- 0

# Remove anything with 0% rel abun
classVer.rel = prune_taxa(taxa_sums(classVer.rel)>0,classVer.rel)

# Create rank abundance of OTUs
barplot(sort(taxa_sums(classVer.rel),TRUE)[1:20]/nsamples(classVer.rel),las=2,cex.axis=.7)

# Create dataframe to use in ggplot
classVer.df <- psmelt(classVer.rel)
classVer.df$Abundance <- classVer.df$Abundance * 100
#classVer.df$Class <- ordered(classVer.df$Class, levels = c("Opitutae","Verrucomicrobiae","OPB35_soil_Lake_Type", "Spartobacteria", "Verrucomicrobia_Incertae_Sedis", "unclassified", "S-BQ2-57_soil_Lake_Type"))
classVer.df$Class <- ordered(classVer.df$Class, levels = c("S-BQ2-57_soil_Lake_Type", "unclassified", "Verrucomicrobia_Incertae_Sedis", "Spartobacteria", "OPB35_soil_Lake_Type", "Verrucomicrobiae", "Opitutae"))
classVer.df$Lake_Type <- ordered(classVer.df$Lake_Type, levels = c("Laurentian","Estuary","Inland"))
classVer.df$Fraction <- ordered(classVer.df$Fraction, levels = c("Particle","Free","Sediment"))
classVer.df$Depth <- ordered(classVer.df$Depth, levels = c("Top","DCM", "Bottom","Sediment"))
classVer.df$Season <- ordered(classVer.df$Season, levels = c("Spring","Summer","Fall"))

Ver.colors <- c(Opitutae = "red2", Verrucomicrobiae = "darkorange", "OPB35_soil_Lake_Type" = "gold", Spartobacteria = "chartreuse3", "Verrucomicrobia_Incertae_Sedis" = "deepskyblue", unclassified = "mediumorchid3", "S-BQ2-57_soil_Lake_Type" = "purple4")

#phylum.colors <- c(Proteobacteria = "deepskyblue", Actinobacteria = "royalblue", Cyanobacteria = "chartreuse3", Bacteroidetes = "darkorange", Planctomycetes = "mediumorchid3", Chloroflexi = "darkgreen", Verrucomicrobia = "purple4", Armatimonadetes = "red", "NPL-UPA2"="#652926", Chlorobi = "cyan2", Gemmatimonadetes = "black", Acidobacteria = "grey26", "Candidate_division_OP3" = "hotpink", Nitrospirae = "blue", "Candidate_division_OD1" = "brown", Spirochaetae = "gold3", "Candidate_division_OP8" = "goldenrod1", Chlamydiae = "violet", Lentisphaerae = "yellow1", "BD1-5" = "chartreuse", "TM6" = "olivedrab", unclassified = "grey", Firmicutes = "blue4", "Candidate_division_WS3"= "magenta", Fibrobacteres = "darkred", "Candidate_division_SR1" = "tan3" )






########### Create dataframe for unstacked bar plots (facet by Lake_Type & Fraction) ###########
classVer.rel.df <- psmelt(classVer.rel)
    ## This takes your entire phyloseq object & combines its constituents into one dataframe
sub.classVer.df <- subset(classVer.df, select = c("Sample","Class", "Abundance","Lake_Type","Fraction"))
    ## These are the factors that you'll be using to make your graph
Lake_Typefrac <- ddply(sub.classVer.df, c("Lake_Type","Fraction", "Class"), summarise,
                   n =  length(Abundance),
                   relabun = mean(Abundance),
                   sd = sd(Abundance),
                   se = sd/sqrt(n) )
    ## This takes one dataframe, applies a bunch of functions to it, and spits out the results
    ## into a new dataframe
    ## ddply ( dataframe.you.wanna.work.with, c("the factors you want to use"),
    ##         summarise (no clue what this adds to the output),
    ##         now all the functions/commands you want to apply

Lake_Typefrac$Class <- ordered(Lake_Typefrac$Class, levels = c("S-BQ2-57_soil_Lake_Type", "unclassified", "Verrucomicrobia_Incertae_Sedis", "Spartobacteria", "OPB35_soil_Lake_Type", "Verrucomicrobiae", "Opitutae"))
Lake_Typefrac$Lake_Type <- ordered(Lake_Typefrac$Lake_Type, levels = c("Laurentian","Estuary","Inland"))
Lake_Typefrac$Fraction <- ordered(Lake_Typefrac$Fraction, levels = c("Particle","Free","Sediment"))

Ver.colors <- c(Opitutae = "red3", Verrucomicrobiae = "darkorange", "OPB35_soil_Lake_Type" = "gold", Spartobacteria = "chartreuse3", "Verrucomicrobia_Incertae_Sedis" = "steelblue1", unclassified = "mediumorchid3", "S-BQ2-57_soil_Lake_Type" = "purple4")


###### GGPLOT IT! ######
tiff("unstackbar.tiff", width = 8, height = 5, units = "in", res = 500)
ggplot(Lake_Typefrac, aes(y=relabun, x = Class, fill = Class)) +
  facet_grid(Fraction ~ Lake_Type) +
  geom_bar(stat="identity", position=position_dodge()) +
  scale_fill_manual(values = Ver.colors, name = "Class") +
  theme_bw() + ggtitle("Verrucomicrobia Classes Above 0.1%") + coord_flip() +
  geom_errorbar(aes(ymin = relabun - se, ymax = relabun + se), width = 0.25) + 
  xlab("Class") + ylab("Mean Relative Abundance (%)") +
  theme(axis.title.x = element_text(face="bold", size=16),
        axis.text.x = element_text(angle=0, colour="black", size=12),
        axis.text.y = element_text(colour="black", size=12),
        axis.title.y = element_text(face="bold", size=16),
        plot.title = element_text(face="bold", size=20),
        legend.title = element_text(size=12, face="bold"),
        legend.text = element_text(size = 12),
        legend.position="none")
dev.off()







########### Create dataframe for unstacked bar plots (facet by Season & Fraction) ###########
classVer.rel.df <- psmelt(classVer.rel)
    ## This takes your entire phyloseq object & combines its constituents into one dataframe
sub.classVer.df <- subset(classVer.df, select = c("Sample","Class", "Abundance","Season", "Fraction"))
    ## These are the factors that you'll be using to make your graph
Lake_Typefrac <- ddply(sub.classVer.df, c("Season", "Class", "Fraction"), summarise,
                   n =  length(Abundance),
                   relabun = mean(Abundance),
                   sd = sd(Abundance),
                   se = sd/sqrt(n) )
    ## This takes one dataframe, applies a bunch of functions to it, and spits out the results
    ## into a new dataframe
    ## ddply ( dataframe.you.wanna.work.with, c("the factors you want to use"),
    ##         summarise (no clue what this adds to the output),
    ##         now all the functions/commands you want to apply

Lake_Typefrac$Season <- ordered(Lake_Typefrac$Season, levels = c("Spring","Summer","Fall"))
Lake_Typefrac$Fraction <- ordered(Lake_Typefrac$Fraction, levels = c("Particle","Free","Sediment"))

Ver.colors <- c(Opitutae = "red3", Verrucomicrobiae = "darkorange", "OPB35_soil_Lake_Type" = "gold", Spartobacteria = "chartreuse3", "Verrucomicrobia_Incertae_Sedis" = "steelblue1", unclassified = "mediumorchid3", "S-BQ2-57_soil_Lake_Type" = "purple4")


###### GGPLOT IT! ######
#tiff("unstackbar.seasonfrac.tiff", width = 8, height = 5, units = "in", res = 500)
ggplot(Lake_Typefrac, aes(y=relabun, x = Class, fill = Class)) +
  facet_grid(Fraction ~ Season) +
  geom_bar(stat="identity", position=position_dodge()) +
  scale_fill_manual(values = Ver.colors, name = "Class") +
  theme_bw() + ggtitle("Verrucomicrobia Classes Above 0.1%") + coord_flip() +
  geom_errorbar(aes(ymin = relabun - se, ymax = relabun + se), width = 0.25) + 
  xlab("Class") + ylab("Mean Relative Abundance (%)") +
  theme(axis.title.x = element_text(face="bold", size=16),
        axis.text.x = element_text(angle=0, colour="black", size=12),
        axis.text.y = element_text(colour="black", size=12),
        axis.title.y = element_text(face="bold", size=16),
        plot.title = element_text(face="bold", size=20),
        legend.title = element_text(size=12, face="bold"),
        legend.text = element_text(size = 12),
        legend.position="none")
#dev.off()




########### Create dataframe for unstacked bar plots (facet by Lake_Type & Season) ###########
classVer.rel.df <- psmelt(classVer.rel)
    ## This takes your entire phyloseq object & combines its constituents into one dataframe
sub.classVer.df <- subset(classVer.df, select = c("Sample","Class", "Abundance","Season", "Lake_Type"))
    ## These are the factors that you'll be using to make your graph
Lake_Typefrac <- ddply(sub.classVer.df, c("Season", "Class", "Lake_Type"), summarise,
                   n =  length(Abundance),
                   relabun = mean(Abundance),
                   sd = sd(Abundance),
                   se = sd/sqrt(n) )
    ## This takes one dataframe, applies a bunch of functions to it, and spits out the results
    ## into a new dataframe
    ## ddply ( dataframe.you.wanna.work.with, c("the factors you want to use"),
    ##         summarise (no clue what this adds to the output),
    ##         now all the functions/commands you want to apply

Lake_Typefrac$Season <- ordered(Lake_Typefrac$Season, levels = c("Spring","Summer","Fall"))
Lake_Typefrac$Lake_Type <- ordered(Lake_Typefrac$Lake_Type, levels = c("Laurentian","Estuary","Inland"))

Ver.colors <- c(Opitutae = "red3", Verrucomicrobiae = "darkorange", "OPB35_soil_Lake_Type" = "gold", Spartobacteria = "chartreuse3", "Verrucomicrobia_Incertae_Sedis" = "steelblue1", unclassified = "mediumorchid3", "S-BQ2-57_soil_Lake_Type" = "purple4")


###### GGPLOT IT! ######
tiff("unstackbar.Lake_Typeseason.tiff", width = 8, height = 5, units = "in", res = 500)
ggplot(Lake_Typefrac, aes(y=relabun, x = Class, fill = Class)) +
  facet_grid(Season ~ Lake_Type) +
  geom_bar(stat="identity", position=position_dodge()) +
  scale_fill_manual(values = Ver.colors, name = "Class") +
  theme_bw() + ggtitle("Verrucomicrobia Classes Above 0.1%") + coord_flip() +
  geom_errorbar(aes(ymin = relabun - se, ymax = relabun + se), width = 0.25) + 
  xlab("Class") + ylab("Mean Relative Abundance (%)") +
  theme(axis.title.x = element_text(face="bold", size=16),
        axis.text.x = element_text(angle=0, colour="black", size=12),
        axis.text.y = element_text(colour="black", size=12),
        axis.title.y = element_text(face="bold", size=16),
        plot.title = element_text(face="bold", size=20),
        legend.title = element_text(size=12, face="bold"),
        legend.text = element_text(size = 12),
        legend.position="none")
dev.off()





########### Create dataframe for unstacked bar plots (facet by Fraction ~ Depth) ###########
classVer.rel.df <- psmelt(classVer.rel)
    ## This takes your entire phyloseq object & combines its constituents into one dataframe
sub.classVer.df <- subset(classVer.df, select = c("Sample","Class", "Abundance","Depth", "Fraction"))
    ## These are the factors that you'll be using to make your graph
Lake_Typefrac <- ddply(sub.classVer.df, c("Depth", "Class", "Fraction"), summarise,
                   n =  length(Abundance),
                   relabun = mean(Abundance),
                   sd = sd(Abundance),
                   se = sd/sqrt(n) )
    ## This takes one dataframe, applies a bunch of functions to it, and spits out the results
    ## into a new dataframe
    ## ddply ( dataframe.you.wanna.work.with, c("the factors you want to use"),
    ##         summarise (no clue what this adds to the output),
    ##         now all the functions/commands you want to apply

Lake_Typefrac$Depth <- ordered(Lake_Typefrac$Depth, levels = c("Top","DCM","Bottom","Sediment"))
Lake_Typefrac$Fraction <- ordered(Lake_Typefrac$Fraction, levels = c("Particle","Free","Sediment"))


Ver.colors <- c(Opitutae = "red3", Verrucomicrobiae = "darkorange", "OPB35_soil_Lake_Type" = "gold", Spartobacteria = "chartreuse3", "Verrucomicrobia_Incertae_Sedis" = "steelblue1", unclassified = "mediumorchid3", "S-BQ2-57_soil_Lake_Type" = "purple4")


###### GGPLOT IT! ######
#tiff("unstackbar.depthfrac.tiff", width = 8, height = 5, units = "in", res = 500)
ggplot(Lake_Typefrac, aes(y=relabun, x = Class, fill = Class)) +
  facet_grid(Fraction ~ Depth) +
  geom_bar(stat="identity", position=position_dodge()) +
  scale_fill_manual(values = Ver.colors, name = "Class") +
  theme_bw() + ggtitle("Verrucomicrobia Classes Above 0.1%") + coord_flip() +
  geom_errorbar(aes(ymin = relabun - se, ymax = relabun + se), width = 0.25) + 
  xlab("Class") + ylab("Mean Relative Abundance (%)") +
  theme(axis.title.x = element_text(face="bold", size=16),
        axis.text.x = element_text(angle=0, colour="black", size=12),
        axis.text.y = element_text(colour="black", size=12),
        axis.title.y = element_text(face="bold", size=16),
        plot.title = element_text(face="bold", size=20),
        legend.title = element_text(size=12, face="bold"),
        legend.text = element_text(size = 12),
        legend.position="none")
#dev.off()





#good_phylum_proteo <-tax_glom(merged_final,taxrank = "Phylum")
goodsamps_phylum <- tax_glom(good_phylum_proteo,taxrank = "Phylum")
goodsamps_phy_melt <- psmelt(goodsamps_phylum)  ##  Melt it into a dataframe 
sub_goodsamps_phy_melt <- subset(goodsamps_phy_melt, select = c("Sample", "ProdLevel", "quadrant", "Phylum","Abundance"))
TOTALS <- ddply(sub_goodsamps_phy_melt, c("Sample"), summarise, ##  Let's get the McMurdie and Holmes Scaled  Total for each sample 
                total   = sum(Abundance))   
sub_phy_melt_totals <- merge(sub_goodsamps_phy_melt, TOTALS, by = "Sample")  ### Merge phylum sums and the total numbers -> so we can calculate the Relative Abundance
sub_phy_melt_totals$RelAbundance <- sub_phy_melt_totals$Abundance/sub_phy_melt_totals$total  ## Calculate the relative abundance
sub_phy_melt_totals$PercentAbund <- sub_phy_melt_totals$RelAbundance * 100  ##  Calculate the Percent Abundance

####  Make a new dataframe with the percent abudance within the entire dataset!
phy_stats <- ddply(sub_phy_melt_totals, c("Phylum"), summarise, 
                   N = length(PercentAbund),
                   PercentPhy = mean(PercentAbund),
                   sd   = sd(PercentAbund),
                   se   = sd / sqrt(N))
abund <- subset(phy_stats,PercentPhy > 0.001)  # Only take the phyla that are more than 0.01% abundant

abund_ordered <- arrange(abund, desc(PercentPhy))

abund$Phylum <- factor(abund$Phylum,levels = c("Bacteroidetes", "Cyanobacteria", "Verrucomicrobia", "Betaproteobacteria", "Actinobacteria", "Planctomycetes",  "Alphaproteobacteria", "Deltaproteobacteria",
                                               "Gammaproteobacteria", "Chloroflexi", "Lentisphaerae", "Chlorobi", "Armatimonadetes", "Firmicutes", "Acidobacteria", "Spirochaetae", "Candidate_division_OD1",
                                               "NPL-UPA2", "Deinococcus-Thermus", "Candidate_division_OP3", "TM6", "Chlamydiae", "Epsilonproteobacteria", "TA18", "Fibrobacteres", "Candidate_division_SR1", 
                                               "Candidate_division_BRC1", "BD1-5", "Gemmatimonadetes", "Fusobacteria", "Candidate_division_WS3", "Tenericutes", "Elusimicrobia", "WCHB1-60", "Deferribacteres", 
                                               "Candidate_division_TM7", "Candidate_division_OP8", "SPOTSOCT00m83", "Thermotogae", "Candidate_division_OP11", "Dictyoglomi", "unclassified"))   

abund$Phylum = with(abund, factor(Phylum, levels = rev(levels(Phylum)))) 

#Relative abundance plot 
abund_plot <- ggplot(abund, aes(y=PercentPhy , x=Phylum))  +
  #geom_boxplot(fill = "magenta4", colour = "black") + 
  geom_bar(stat="identity", position=position_dodge(),  fill = "magenta4", colour = "black") +
  theme_bw() + ggtitle("Phyla Above 0.1% in All Samples") +
  xlab("Phylum") + ylab("Mean Relative Abundance (%)") +
  geom_errorbar(aes(ymin = PercentPhy -se, ymax = PercentPhy +se), width = 0.25) + coord_flip() +
  theme(axis.title.x = element_text(face="bold", size=16),
        axis.text.x = element_text(angle=0, colour = "black", size=14),
        axis.text.y = element_text(colour = "black", size=14),
        axis.title.y = element_text(face="bold", size=16),
        plot.title = element_text(face="bold", size = 20),
        legend.title = element_text(size=12, face="bold"),
        legend.text = element_text(size = 12),
        legend.position="none"); abund_plot


```


### DESeq
```{r deseq, echo=FALSE}
# DESeq bible: https://www.bioconductor.org/packages/3.3/bioc/vignettes/DESeq/inst/doc/DESeq.pdf
# Phyloseq to DESeq2: http://joey711.github.io/phyloseq-extensions/DESeq2.html
# Phyloseq to DESeq: http://joey711.github.io/phyloseq-extensions/DESeq.html


### DESeq2 ###
# Prepare phyloseq object
Vdeseq <- subset_taxa(dupmerge.subset, Phylum == "Verrucomicrobia")
  # We want to use the RAW COUNTS
pick.Ver<- colnames(otu_table(Ver.rel))
Vdeseq.pick <- prune_taxa(names(taxa_sums(Vdeseq)) %in% pick.Ver, x=Vdeseq)
  # 208 OTUs
  # Subset out the verruco taxa that we've been using so far
  # When we calculate Verruco relative abundance, we remove a few hundred super low abundance OTUs
Ver1 <- otu_table(Vdeseq.pick) + 1
Ver.samp <- sample_data(Vdeseq.pick)
Ver.tax <- tax_table(Vdeseq.pick)
Ver1 <- merge_phyloseq(Ver1, Ver.samp, Ver.tax)




# Editing phyloseq object because there can be no 0's
#Ver1 <- otu_table(Ver) + 1
#Ver.samp <- sample_data(Ver)
#Ver.tax <- tax_table(Ver)
#Ver1 <- merge_phyloseq(Ver1, Ver.samp, Ver.tax)



#test <- phyloseq_to_deseq2(Ver1, design= ~Lake_Type+Season+Source)
#test.deseq <- DESeq(test, test="Wald", fitType="local")
#test.res <- results(test.deseq, cooksCutoff = FALSE)
#test.res.alpha <- results(test.deseq, cooksCutoff=F, alpha=0.05)

#test.deseq.out <- DESeq(test, test="Wald", fitType="local", minReplicatesForReplace=Inf)
#test.res.out <- results(test.deseq.out, cooksCutoff = FALSE)
#test.res.out.alpha <- results(test.deseq.out, cooksCutoff=F, alpha=0.05)


# Let's look at Source (water vs. sed)
source <- deSEQ(data=Ver1, valuetest= ~Lake_Type+Season+Source)
  # I left out Depth & Fraction because those all take into account water vs. sediment

### Separate sediment & water samples
sed <- which(source$log2FoldChange < 0)
sed.otus <- rownames(source[sed,])
water <- which(source$log2FoldChange > 0)
water.otus <- rownames(source[water,])

deseq.source <- ggplot(source, aes(x=Species, y=log2FoldChange, color=Class)) + geom_point() +
  scale_color_brewer(palette="Set1") +
  ggtitle("Sediment (Top) vs. Water (Bottom)")



# Now let's isolate Verrucos in water
Ver.water <- subset_samples(Ver1, Depth != "Sediment")
Ver.water <- subset_samples(Ver.water, Depth != "DCM")
  # We want binary, so remove DCM too
watercol <- deSEQ(data=Ver.water, valuetest= ~Lake_Type+Season+Depth, alpha=0.05)

### Separate sediment & water samples
water.bot <- which(watercol$log2FoldChange < 0)
water.bot.otus <- rownames(watercol[water.bot,])
water.top <- which(watercol$log2FoldChange > 0)
water.top.otus <- rownames(watercol[water.top,])

deseq.watercol<- ggplot(watercol, aes(x=Species, y=log2FoldChange, color=Class)) + geom_point() +
  scale_color_brewer (palette="Set1") +
  ggtitle("Top (Top) vs. Bottom (Bottom)")


# Verrucos in Bottom vs. Sediment
Ver.bs <- subset_samples(Ver1, Depth != "Top")
Ver.bs <- subset_samples(Ver.bs, Depth != "DCM")
Vbs <- deSEQ(data=Ver.bs, valuetest= ~Lake_Type+Season+Depth, alpha=0.05)

### Separate sediment & water samples
bs.bot <- which(Vbs$log2FoldChange < 0)
bs.bot.otus <- rownames(Vbs[bs.bot,])
bs.sed <- which(Vbs$log2FoldChange > 0)
bs.sed.otus <- rownames(Vbs[bs.sed,])

deseq.Vbs <- ggplot(Vbs, aes(x=Species, y=log2FoldChange, color=Class)) + geom_point() +
  scale_color_brewer(palette="Set1") +
  ggtitle("Sediment (Top) vs. Bottom (Bottom)")



# Verrucos in Sediment vs. Top
Ver.st <- subset_samples(Ver1, Depth != "Bottom")
Ver.st <- subset_samples(Ver.st, Depth != "DCM")
Vst <- deSEQ(data=Ver.st, valuetest= ~Lake_Type+Season+Depth, alpha=0.05)

### Separate sediment & water samples
st.sed <- which(Vst$log2FoldChange < 0)
st.sed.otus <- rownames(Vst[st.sed,])
st.top <- which(Vst$log2FoldChange > 0)
st.top.otus <- rownames(Vst[st.top,])

deseq.Vst <- ggplot(Vst, aes(x=Species, y=log2FoldChange, color=Class)) + geom_point() +
  scale_color_brewer(palette="Set1") +
  ggtitle("Top (Top) vs. Sediment (Bottom)")


sed.otus1 <- intersect(water.bot.otus, bs.sed.otus)
  # Which OTUs that prefer bottom, also prefer sed
sed.otus2 <- intersect(water.top.otus, st.sed.otus)
  # Which OTUs that prefer top, also prefer sed
sed.otus <- c(sed.otus1, sed.otus2)
  ## OTUs that are overrepresented in Sediment
bot.otus <- intersect(water.bot.otus, bs.bot.otus)
  ## OTUs that are overrepresented in Bottom
top.otus <- intersect(water.top.otus, st.top.otus)
  ## OTUs that are overrepresented in Top



tiff("deseq.depth.tiff", width = 14, height = 10, units = "in", res = 500)
grid.newpage()
pushViewport(viewport(layout=(grid.layout(2,2, widths=c(1,1), heights=c(1.5,1.5)))))
print(deseq.source, vp=viewport(layout.pos.row=1, layout.pos.col=1))
print(deseq.watercol, vp=viewport(layout.pos.row=1, layout.pos.col=2))
print(deseq.Vst, vp=viewport(layout.pos.row=2, layout.pos.col=1))
print(deseq.Vbs, vp=viewport(layout.pos.row=2, layout.pos.col=2))
dev.off()




### Lake_Type
# Now let's isolate Verrucos in Estuary & Laurentian
V.el <- subset_samples(Ver1, Lake_Type != "Inland")
Vel <- deSEQ(data=V.el, valuetest= ~Depth+Season+Lake_Type, alpha=0.05)

### Separate sediment & water samples
Vel.est <- which(Vel$log2FoldChange < 0)
Vel.est.otus <- rownames(Vel[Vel.est,])
Vel.lau <- which(Vel$log2FoldChange > 0)
Vel.lau.otus <- rownames(Vel[Vel.lau,])

deseq.Vel<- ggplot(Vel, aes(x=Species, y=log2FoldChange, color=Class)) + geom_point() +
  scale_color_brewer(palette="Set1") +
  ggtitle("Laurentian (Top) vs. Estuary (Bottom)")



# Now let's isolate Verrucos in Estuary & Inland
V.ei <- subset_samples(Ver1, Lake_Type != "Laurentian")
Vei <- deSEQ(data=V.ei, valuetest= ~Depth+Season+Lake_Type, alpha=0.05)

### Separate sediment & water samples
Vei.est <- which(Vei$log2FoldChange < 0)
Vei.est.otus <- rownames(Vei[Vel.est,])
Vei.inl <- which(Vei$log2FoldChange > 0)
Vei.inl.otus <- rownames(Vei[Vei.inl,])

deseq.Vei<- ggplot(Vei, aes(x=Species, y=log2FoldChange, color=Class)) + geom_point() +
  scale_color_brewer(palette="Set1") +
  ggtitle("Inland (Top) vs. Estuary (Bottom)")




# Now let's isolate Verrucos in Laurentian & Inland
V.li <- subset_samples(Ver1, Lake_Type != "Estuary")
Vli <- deSEQ(data=V.li, valuetest= ~Depth+Season+Lake_Type, alpha=0.05)

### Separate sediment & water samples
Vli.inl <- which(Vli$log2FoldChange < 0)
Vli.inl.otus <- rownames(Vli[Vli.inl,])
Vli.lau <- which(Vel$log2FoldChange > 0)
Vli.lau.otus <- rownames(Vli[Vli.lau,])

deseq.Vli<- ggplot(Vli, aes(x=Species, y=log2FoldChange, color=Class)) + geom_point() +
  scale_color_brewer(palette="Set1") +
  ggtitle("Laurentian (Top) vs. Inland (Bottom)")


### Lake_Type Otus ###
est.otus <- intersect(Vel.est.otus, Vei.est.otus)
inl.otus <- intersect(Vei.inl.otus, Vli.inl.otus)
lau.otus <- intersect(Vel.lau.otus, Vli.lau.otus)

tiff("deseq.Lake_Type.tiff", width = 20, height = 10, units = "in", res = 500)
grid.newpage()
pushViewport(viewport(layout=(grid.layout(1,3, widths=c(1.25,1.25,1.25), heights=c(1.5,1.5)))))
print(deseq.Vel, vp=viewport(layout.pos.row=1, layout.pos.col=1))
print(deseq.Vei, vp=viewport(layout.pos.row=1, layout.pos.col=2))
print(deseq.Vli, vp=viewport(layout.pos.row=1, layout.pos.col=3))
dev.off()





####### Fraction
# Now let's isolate Verrucos in Free & Particle (fractions, so no sediment but DCM is included)
V.frac <- subset_samples(Ver1, Fraction != "Sediment")
Vfrac <- deSEQ(data=V.frac, valuetest= ~Depth+Lake_Type+Season+Fraction, alpha=0.05)

### Separate sediment & water samples
Vfrac.free <- which(Vfrac$log2FoldChange < 0)
free.otus <- rownames(Vfrac[Vfrac.free,])
Vfrac.part <- which(Vfrac$log2FoldChange > 0)
part.otus <- rownames(Vfrac[Vfrac.part,])

tiff("deseq.fraction.tiff", width = 7, height = 10, units = "in", res = 500)
ggplot(Vfrac, aes(x=Species, y=log2FoldChange, color=Class)) + geom_point() +
  scale_color_brewer(palette="Set1") +
  ggtitle("Particle (Top) vs. Free (Bottom)")
dev.off()








####### Season
# Now let's isolate Verrucos in Spring & Summer
V.ss <- subset_samples(Ver1, Season != "Fall")
Vss <- deSEQ(data=V.ss, valuetest= ~Depth+Lake_Type+Season, alpha=0.05)

### Separate sediment & water samples
Vss.sp <- which(Vss$log2FoldChange < 0)
Vss.sp.otus <- rownames(Vss[Vss.sp,])
Vss.su <- which(Vss$log2FoldChange > 0)
Vss.su.otus <- rownames(Vss[Vss.su,])

deseq.Vss <- ggplot(Vss, aes(x=Species, y=log2FoldChange, color=Class)) + geom_point() +
  scale_color_brewer(palette="Set1") +
  ggtitle("Season: Summer (Top) vs. Spring (Bottom)")






# Now let's isolate Verrucos in Spring & Fall
V.spf <- subset_samples(Ver1, Season != "Summer")
Vspf <- deSEQ(data=V.spf, valuetest= ~Depth+Lake_Type+Season, alpha=0.05)

### Separate sediment & water samples
Vspf.fa <- which(Vspf$log2FoldChange < 0)
Vspf.fa.otus <- rownames(Vspf[Vspf.fa,])
Vspf.sp <- which(Vspf$log2FoldChange > 0)
Vspf.sp.otus <- rownames(Vspf[Vspf.sp,])

deseq.Vspf <- ggplot(Vspf, aes(x=Species, y=log2FoldChange, color=Class)) + geom_point() +
  scale_color_brewer(palette="Set1") +
  ggtitle("Season: Spring (Top) vs. Fall (Bottom)")








# Now let's isolate Verrucos in Summer & Fall
V.suf <- subset_samples(Ver1, Season != "Spring")
Vsuf <- deSEQ(data=V.suf, valuetest= ~Depth+Lake_Type+Season, alpha=0.05)

### Separate sediment & water samples
Vsuf.fa <- which(Vsuf$log2FoldChange < 0)
Vsuf.fa.otus <- rownames(Vsuf[Vsuf.fa,])
Vsuf.su <- which(Vsuf$log2FoldChange > 0)
Vsuf.su.otus <- rownames(Vsuf[Vsuf.su,])

deseq.Vsuf <- ggplot(Vsuf, aes(x=Species, y=log2FoldChange, color=Class)) + geom_point() +
  scale_color_brewer(palette="Set1") +
  ggtitle("Season: Summer (Top) vs. Fall (Bottom)")



sp.otus <- intersect(Vss.sp.otus,Vspf.sp.otus)
su.otus <- intersect(Vss.su.otus,Vsuf.su.otus)
fa.otus <- intersect(Vspf.fa.otus,Vsuf.fa.otus)



tiff("deseq.season.tiff", width = 20, height = 10, units = "in", res = 500)
grid.newpage()
pushViewport(viewport(layout=(grid.layout(1,3, widths=c(1.25,1.25,1.25), heights=c(1.5,1.5)))))
print(deseq.Vss, vp=viewport(layout.pos.row=1, layout.pos.col=1))
print(deseq.Vspf, vp=viewport(layout.pos.row=1, layout.pos.col=2))
print(deseq.Vsuf, vp=viewport(layout.pos.row=1, layout.pos.col=3))
dev.off()
```



### Create ecoIDs
```{r ecoID}
#http://stackoverflow.com/questions/10298662/find-elements-not-in-smaller-character-vector-list-but-in-big-list
Ver.otus <- rownames(tax_table(Ver1))

Lake_Type.label <- rownames(tax_table(Ver1))
Lake_Type.label[Lake_Type.label %in% est.otus] <- paste("E", Lake_Type.label[Lake_Type.label %in% est.otus])
Lake_Type.label[Lake_Type.label %in% inl.otus] <- paste("I", Lake_Type.label[Lake_Type.label %in% inl.otus])
Lake_Type.label[Lake_Type.label %in% lau.otus] <- paste("L", Lake_Type.label[Lake_Type.label %in% lau.otus])
# 57 labels; 1265 OTUs have no preference

Depth.label <- rownames(tax_table(Ver1))
Depth.label[Depth.label %in% bot.otus] <- paste("B", Depth.label[Depth.label %in% bot.otus])
Depth.label[Depth.label %in% sed.otus] <- paste("S", Depth.label[Depth.label %in% sed.otus])
Depth.label[Depth.label %in% top.otus] <- paste("T", Depth.label[Depth.label %in% top.otus])
# 107 labels; 1215 OTUs have no preference

Fraction.label <- rownames(tax_table(Ver1))
Fraction.label[Fraction.label %in% free.otus] <- paste("F", Fraction.label[Fraction.label %in%
                                                                             free.otus])
Fraction.label[Fraction.label %in% part.otus] <- paste("P", Fraction.label[Fraction.label %in%
                                                                             part.otus])
# 80 labels; 1242 OTUs have no preference

Season.label <- rownames(tax_table(Ver1))
Season.label[Season.label %in% sp.otus] <- paste("C_", Season.label[Season.label %in% sp.otus])
Season.label[Season.label %in% su.otus] <- paste("V_", Season.label[Season.label %in% su.otus])
Season.label[Season.label %in% fa.otus] <- paste("A_", Season.label[Season.label %in% fa.otus])
# 56 labels; 1266 OTUs have no preference

Lake_Type.split <- data.frame(strsplit(Lake_Type.label, " "), stringsAsFactors=FALSE)
colnames(Lake_Type.split) <- Ver.otus
depth.split <- data.frame(strsplit(Depth.label, " "), stringsAsFactors=FALSE)
colnames(depth.split) <- Ver.otus
fraction.split <- data.frame(strsplit(Fraction.label, " "), stringsAsFactors=FALSE)
colnames(fraction.split) <- Ver.otus
season.split <- data.frame(strsplit(Season.label, " "), stringsAsFactors=FALSE)
colnames(season.split) <- Ver.otus
Lake_Type.split[3,] <- depth.split[1,]
Lake_Type.split[4,] <- fraction.split[1,]
Lake_Type.split[5,] <- season.split[1,]
Lake_Type.split <- Lake_Type.split[-2,]
# Now to rename the OTUs with no preference in one of the ecological factors
nopref <- which(substr(Lake_Type.split[1,],1,1) == "O")
Lake_Type.split[1,nopref] <- 0
nopref <- which(substr(Lake_Type.split[2,],1,1) == "O")
Lake_Type.split[2,nopref] <- 0
nopref <- which(substr(Lake_Type.split[3,],1,1) == "O")
Lake_Type.split[3,nopref] <- 0
nopref <- which(substr(Lake_Type.split[4,],1,1) == "O")
Lake_Type.split[4,nopref] <- "0_"
Lake_Type.split[5,] <- paste(Lake_Type.split[2,], Lake_Type.split[1,], Lake_Type.split[3,], Lake_Type.split[4,],
                         colnames(Lake_Type.split))
write.csv(Lake_Type.split[5,], file="Ecology_Otu_sig", row.names=FALSE, quote=FALSE)
```

### Create updated phyloseq object with EcoID
```{r ecoid.phyloseq}
##### Create new phyloseq object with EcoIDs
eco.label<- read.csv("Ecology_Otu_sig", header=F)
eco.label <- t(eco.label)
otu <- otu_table(Ver1)
samp <- sample_data(Ver1)
tax <- data.frame(tax_table(Ver1))
tax$EcoID <- eco.label
Lake_Type.split <- t(Lake_Type.split)
tax$Pref.Depth <- Lake_Type.split[,3]
tax$Pref.Lake_Type <- Lake_Type.split[,1]
tax$Pref.Fraction <- Lake_Type.split[,4]
tax$Pref.Season <- Lake_Type.split[,5]
tax <- as.matrix(tax)
tax <- tax_table(tax)
eco.ver <- merge_phyloseq(otu, samp, tax)
newick <- read.newick(file="ver.tre")
newick_phy <- phy_tree(newick)
#test <- import_mothur(mothur_tree_file=newick)
Ver.tree<-merge_phyloseq(eco.ver,newick_phy)
#plot_tree(Ver.tree, color="Class")
```

### Create labels for IToL!
```{r, itol}
### IToL Label Functions
color.id <- function(x) {
  ifelse(grepl("OPB35_soil_group",x),"#e41a1c",
  ifelse(grepl("Opitutae",x), "#377eb8",
  ifelse(grepl("Spartobacteria",x), "#4daf4a",
  ifelse(grepl("S-BQ2-57_soil_group",x), "#984ea3",
  ifelse(grepl("Verrucomicrobia_Incertae_Sedis",x), "#ff7f00",
  ifelse(grepl("Verrucomicrobiae",x), "#ffd92f",
  ifelse(grepl("unclassified",x), "#f1a3a9",
  ifelse(grepl("UA11",x), "#fff5eb",
         NA))))))))
}

depth.id <- function(x) {
  depth <- substr(x$EcoID,1,1)
  ifelse(grepl("T",depth), "1,-1,-1",
  ifelse(grepl("B",depth), "-1,1,-1",
  ifelse(grepl("S",depth), "-1,-1,1",
  ifelse(grepl("0",depth), "-1,-1,-1",
  NA))))
}

Lake_Type.id <- function(x) {
  Lake_Type <- substr(x$EcoID,2,2)
  ifelse(grepl("L",Lake_Type), "1,-1,-1",
  ifelse(grepl("E",Lake_Type), "-1,1,-1",
  ifelse(grepl("I",Lake_Type), "-1,-1,1",
  ifelse(grepl("0", Lake_Type), "-1,-1,-1",
  NA))))
}

fraction.id <- function(x) {
  fraction <- substr(x$EcoID,3,3)
  ifelse(grepl("P",fraction), "1,-1",
  ifelse(grepl("F",fraction), "-1,1",
  ifelse(grepl("0",fraction), "-1,-1",
  NA)))
}

season.id <- function(x) {
  season <- substr(x$EcoID,4,4)
  ifelse(grepl("C",season), "1,-1,-1",
  ifelse(grepl("V",season), "-1,1,-1",
  ifelse(grepl("A",season), "-1,-1,1",
  ifelse(grepl("0", season), "-1,-1,-1",
  NA))))
}

### Create IToL COLOR labels
all <- data.frame(tax_table(Ver.tree))
color.all <- data.frame(all$Species)
color.all$Class <- all$Class
color.all$color <- color.id(color.all$Class)
color.all$label <- paste(color.all$all.Species,"range",color.all$color,color.all$Class)
write.csv(color.all$label,file="sig.ver.color",row.names=FALSE,quote=FALSE)

### Create IToL ECOID Labels
ecoid.all <- data.frame(all$Species)
ecoid.all$EcoID <- all$EcoID
depth.itol <- depth.id(ecoid.all)
ecoid.all$depth.label <- depth.itol
Lake_Type.itol <- Lake_Type.id(ecoid.all)
ecoid.all$Lake_Type.label <- Lake_Type.itol
fraction.itol <- fraction.id(ecoid.all)
ecoid.all$fraction.label <- fraction.itol
season.itol <- season.id(ecoid.all)
ecoid.all$season.label <- season.itol
ecoid.all$itol.label <- paste(ecoid.all$all.Species, ecoid.all$depth.label, ecoid.all$Lake_Type.label,
                              ecoid.all$fraction.label, ecoid.all$season.label)
write.csv(ecoid.all$itol.label,"EcoID.IToL.label.sig", row.names=FALSE, quote=FALSE)
```


### Subset only significant EcoID Verrucos
```{r itol.sig}
nosig <- which(substr(ecoid.all$EcoID,1,4)=="0000")
nosig.otus <- as.character(ecoid.all$all.Species[nosig])
sig.otus <- Ver.otus[!Ver.otus %in% nosig.otus]
only.sig <- prune_taxa(sig.otus, Ver.tree)
write.tree(phy_tree(only.sig), file="ver.sig.tre")

```


### Create dataset for ITOL for top 50 Verruco OTUS
```{r ITOL.100}
# Subset out top 100 most abundant Verruco OTUs
top50ver = sort(taxa_sums(Ver), TRUE)[1:50]
print(top50ver)
top50ver.names<-names(top50ver)

Ver.50 <- prune_taxa(mostabund.names, Ver)
#write.tree(phy_tree(Ver.50), file="ver.50.tre")
```


### Create dataset for ITOL for top 100 Verruco OTUS
```{r ITOL.100}
# Subset out top 100 most abundant Verruco OTUs
top100ver = sort(taxa_sums(Ver), TRUE)[1:100]
print(top100ver)
top100ver.names<-names(top100ver)

Ver.100 <- prune_taxa(mostabund.names, Ver)
#write.tree(phy_tree(Ver.100), file="ver.100.tre")
```


### Check Sig Verruco Abundance
```{r sig.itol.abund}
remain.tree <- read.newick(file="ver.sig.tre")
remain.otus <- remain.tree$tip.label

#remove.tree <- read.newick(file="ver.remove.sig.01.tre")
#remove.otus <- remove.tree$tip.label

#remain.unique <- setdiff(remain.otus, remove.otus)
  # remain.otus that are not in remove.otus
#remove.unique <- setdiff(remove.otus, remain.otus)
  # remove.otus that are not in remain.otus

remain.phy <- prune_taxa(names(taxa_sums(Vdeseq.pick)) %in% remain.otus, x=Vdeseq.pick)
remain.phy <- merge_phyloseq(remain.phy, remain.tree)
#write.tree(phy_tree(refit.phy), file="refit.tre")
remain.rel <- transform_sample_counts(remain.phy, function(x){x/sum(x)})
summary(taxa_sums(remain.rel))
# Create rank abundance of OTUs
barplot(sort(taxa_sums(remain.rel),TRUE)/nsamples(remain.rel),las=2,cex.axis=.7)
# Create rank abundance of OTUs
barplot(sort(taxa_sums(remain.rel),TRUE)[1:60]/nsamples(remain.rel),las=2,cex.axis=.7)



remain.relabund <- prune_taxa(names(taxa_sums(Ver.rel)) %in% remain.otus, x=Ver.rel)

#remain.unique.phy <- prune_taxa(names(taxa_sums(Vdeseq.pick)) %in% remain.unique, x=Vdeseq.pick)

#remove.phy <- prune_taxa(names(taxa_sums(Vdeseq.pick)) %in% remove.otus, x=Vdeseq.pick)
#remove.unique.phy <- prune_taxa(names(taxa_sums(Vdeseq.pick)) %in% remove.unique, x=Vdeseq.pick)
```





### Geiger
```{r geiger}
# https://github.com/mwpennell/geiger-v2

ver.sig.tre <- read.newick(file="ver.sig.tre")

eco.label <- read.csv("Ecology_Otu_sig", header=F)
eco.label <- t(eco.label)
eco.label <- data.frame(eco.label)
eco.label$OTU <- substr(eco.label[,1],6,14)
eco.label$Depth <- substr(eco.label[,1],1,1)
eco.label$Lake_Type <- substr(eco.label[,1],2,2)
eco.label$Fraction <- substr(eco.label[,1],3,3)
eco.label$Season <- substr(eco.label[,1],4,4)

# Remove non-significant Verruco OTUs
nosig <- setdiff(eco.label$OTU, ver.sig.tre$tip.label)
nosig.loc <- pmatch(nosig, eco.label$OTU)
eco.label <- eco.label[-nosig.loc,]


# Pull out only significant depth preferences
eco.depth <- data.frame(eco.label$Depth, row.names=eco.label$OTU)
pref <- which(eco.depth != "0")
sig.depth <- data.frame(eco.depth[pref,1], row.names=row.names(eco.depth)[pref])
nopref <- which(eco.depth == "0")
nopref<- row.names(eco.depth)[nopref]
depth.tre <- drop.tip(ver.sig.tre, nopref)

# Test for phylogenetic signal in depth affinity
optimal <- fitDiscrete(depth.tre, sig.depth, treeTransform="lambda")
phy.Lambda0 <- rescale(depth.tre, "lambda", 0)
zero <- fitDiscrete(phy.Lambda0, sig.depth)
d.lambda=2*(optimal$opt$lnL-zero$opt$lnL)
p.lambda <- pchisq(d.lambda, 1, lower.tail=F)
p.lambda
  # p-value = 0.1857386



# Pull out only significant lake type preferences
eco.lake <- data.frame(eco.label$Lake_Type, row.names=eco.label$OTU)
pref <- which(eco.lake != "0")
sig.lake <- data.frame(eco.lake[pref,1], row.names=row.names(eco.lake)[pref])
nopref <- which(eco.lake == "0")
nopref<- row.names(eco.lake)[nopref]
lake.tre <- drop.tip(ver.sig.tre, nopref)

# Test for phylogenetic signal in lake affinity
optimal <- fitDiscrete(lake.tre, sig.lake, treeTransform="lambda")
phy.Lambda0 <- rescale(lake.tre, "lambda", 0)
zero <- fitDiscrete(phy.Lambda0, sig.lake)
d.lambda=2*(optimal$opt$lnL-zero$opt$lnL)
p.lambda <- pchisq(d.lambda, 1, lower.tail=F)
p.lambda
  # p-value = 0.0286439


# Pull out only significant fraction preferences
eco.frac <- data.frame(eco.label$Fraction, row.names=eco.label$OTU)
pref <- which(eco.frac != "0")
sig.frac <- data.frame(eco.frac[pref,1], row.names=row.names(eco.frac)[pref])
nopref <- which(eco.frac == "0")
nopref<- row.names(eco.frac)[nopref]
frac.tre <- drop.tip(ver.sig.tre, nopref)

# Test for phylogenetic signal in lake affinity
optimal <- fitDiscrete(frac.tre, sig.frac, treeTransform="lambda")
phy.Lambda0 <- rescale(frac.tre, "lambda", 0)
zero <- fitDiscrete(phy.Lambda0, sig.frac)
d.lambda=2*(optimal$opt$lnL-zero$opt$lnL)
p.lambda <- pchisq(d.lambda, 1, lower.tail=F)
p.lambda
  # p-value = 7.514368e-11



# Pull out only significant season preferences
eco.season <- data.frame(eco.label$Season, row.names=eco.label$OTU)
pref <- which(eco.season != "0")
sig.season <- data.frame(eco.season[pref,1], row.names=row.names(eco.season)[pref])
nopref <- which(eco.season == "0")
nopref<- row.names(eco.season)[nopref]
season.tre <- drop.tip(ver.sig.tre, nopref)

# Test for phylogenetic signal in season affinity
optimal <- fitDiscrete(season.tre, sig.season, treeTransform="lambda")
phy.Lambda0 <- rescale(season.tre, "lambda", 0)
zero <- fitDiscrete(phy.Lambda0, sig.season)
d.lambda=2*(optimal$opt$lnL-zero$opt$lnL)
p.lambda <- pchisq(d.lambda, 1, lower.tail=F)
p.lambda
  # p-value = 2.492829e-05

##### Lake Type, Fraction, & Season have significant phylogenetic signals
##### Depth does not have significant phylogenetic signals

```










### ConsenTRAIT
```{r consenTRAIT}
### Prep labels
labels<- read.csv("EcoID.IToL.label.refit", header=F)
for(i in 2:12) {
  labels[,i] <- ifelse(grepl("-1",labels[,i]), "0", "1")
  print(labels)
}

for(i in 2:12) {
  labels[,i] <- as.numeric(labels[,i])
  print(labels)
}






############## Below is ConsenTRAIT script ##############



library("data.table")
library("adephylo")
library("ape")

args <- commandArgs(TRUE)


tree_all <- read.newick(file="refit.tre")
table <- labels

#loading Newick tree (multitree) - replace to read.nexus if using nexus tress
#tree_all = read.tree(args[1],keep.multi = TRUE)
#Loading trait table w. no headers
#table = read.table("table", sep = ",", header=FALSE)



#Starting script
Mean_all<-matrix(nrow=ncol(table)-1,ncol=100)

for (m in 1:length(tree_all)) {#loop through all trees

  tree<-tree_all
# testing if table and tree contain the same entries - else drop tips
  z<-subset(tree$tip.label,!(tree$tip.label %in% table[,1]))
  if (length(z) > 0) {
    drop.tip(tree,z)
  }

#rooting tree with first taxon - change if different root
  root_tree<-root(tree,1,resolve.root=T)
#replacing negative branch lengths - e.g., from PHYLIP
  root_tree$edge.length[root_tree$edge.length <= 0] =  0.00001
  subtree<-subtrees(root_tree, wait=FALSE)


  cluster_mean<-numeric(length=0)

# loop through all traits
  for (j in 2:ncol(table)) {
     print(c("Analyzing",j,"!!!!!!!!!!!!!!!!!!!!!!!"))
     #Loading trait table
     table_tmp<-table[,c(1,j)]
     colnames(table_tmp)[1]<-"ID";
     colnames(table_tmp)[2]<-"Trait";
     
# removing all entries not in tree 
    table_tmp2<-data.table(table_tmp)
     setkey(table_tmp2,ID)
     table2<-table_tmp2[intersect(table_tmp2$ID,root_tree$tip.label)]
     setkey(table2,ID)

     #initializing result vectors and file names
     positives<-vector(mode="list",length=0)
     cluster_size<-numeric(length=0)
     cluster_size_file<-paste("R_cluster_size_",j,".txt",sep="")

     cluster_dist<-numeric(length=0)
     cluster_dist_file<-paste("R_cluster_dist_",j,".txt",sep="")

#initalizing files
     if (m == 1) {
         cat(c("trait","tree","distance","cluster_size"), file = cluster_size_file, sep = "\t", fill = FALSE, labels = NULL,append = FALSE)
         cat("\n", file = cluster_size_file, fill = FALSE, labels = NULL,append = TRUE)

         cat(c("trait","tree","distance"), file = cluster_dist_file, sep = "\t", fill = FALSE, labels = NULL,append = FALSE)
         cat("\n", file = cluster_dist_file, fill = FALSE, labels = NULL,append = TRUE)
     }






     #loop through all subtrees and determining if any subtrees have >90% positives
     for (i in 1:length(subtree)){
       tip_names<-subtree[[1]]$tip.label
       if (mean(table2[tip_names]$Trait) > 0.9 ) {#change this value if you want a new threshold
        match_test<-match(tip_names,positives)
        if (all(is.na(match_test))) {
            positives<-c(positives,tip_names)
            cluster_dist<-distRoot(subtree[[i]],tip_names, method=c("p"))
            cluster_size<-c(cluster_size,mean(cluster_dist))

            # printing to files###
            cat(c(j,m,mean(cluster_dist),length(cluster_dist)), file = cluster_size_file, sep = "\t", fill = FALSE, labels = NULL,append = TRUE)
            cat("\n", file = cluster_size_file, fill = FALSE, labels = NULL,append = TRUE)

            cat(j,m,cluster_dist, file = cluster_dist_file, sep = "\t", fill = FALSE, labels = NULL,append = TRUE)
            cat("\n", file = cluster_dist_file, fill = FALSE, labels = NULL,append = TRUE)


            #print(cluster_dist)
        }
        else if (any(is.na(match_test))) {
            print("some NAs - something is weird")
        }
        else {
            #print(tip_names)
            #print("found cluster before")
        }
      }
    }


    ##### find singletons ######
    a<-table2[table2$Trait == 1,][,ID]
    g<-as.character(a)

    singletons_names = setdiff(g,positives)
    if (length(singletons_names) > 0) {
       for (h in 1:length(singletons_names)){
           # weigh singletons with half
           singleton_edges = 0.5*root_tree$edge.length[which.edge(root_tree,singletons_names[h])] #here we use half the distance for singletons
           cluster_size<-c(cluster_size,singleton_edges)

           cat(c(j,m,singleton_edges,1), file = cluster_size_file, sep = "\t", fill = FALSE, labels = NULL,append = TRUE)
           cat("\n", file = cluster_size_file, sep = "\t", fill = FALSE, labels = NULL,append = TRUE)
       }
    
    }
  Mean_all[j-1,m] = mean(cluster_size)
  }
  

}
#output file
write.table(Mean_all,"Mean_all_bootstrap.txt", sep = "\t")



```






### Extra Functions
```{r}
################# FUNCTION TO FIND NAS IN A DATAFRAME #################
nacols <- function(df) {
  colnames(df)[unlist(lapply(df, function(x) any(is.na(x))))]
}

deSEQ <- function(data, valuetest, cutoff = 0, alpha = 0.01){
  data_pruned <- prune_taxa(taxa_sums(data) > cutoff, data)
  de_data <- phyloseq_to_deseq2(data_pruned, valuetest)
  de_data2 <- DESeq(de_data, test="Wald", fitType="local")
  res_data <- results(de_data2, cooksCutoff = FALSE)
  sig_data <- res_data[which(res_data$padj < alpha), ]
  sigtab_sherm <- cbind(as(sig_data, "data.frame"), as(tax_table(data_pruned)[rownames(sig_data), ], "matrix"))
} 

plot_deSEQ <- function(deSEQdata, title){
  y = tapply(deSEQdata$log2FoldChange, deSEQdata$Genus, function(x) max(x))
  y = sort(y, TRUE)
  deSEQdata$Genus = factor(as.character(deSEQdata$Genus), levels=names(y))
  ggplot(deSEQdata, aes(x=Genus, y=log2FoldChange, color=Phylum)) + 
    geom_point(size=6) +  ggtitle(title) +  
    scale_color_manual(values = phylum.colors,name="Phylum") +
    scale_y_continuous(breaks=seq(-15,15,1)) + 
    theme(axis.title.x = element_text(face="bold", size=16),
          axis.text.x = element_text(angle=30, colour = "black", vjust=1, hjust = 1, size=12),
          axis.text.y = element_text(colour = "black", size=16),
          axis.title.y = element_text(face="bold", size=16),
          plot.title = element_text(face="bold", size = 20),
          legend.title = element_text(size=12, face="bold"),
          legend.text = element_text(size = 12),
          strip.background = element_rect(colour="black"),
          legend.position="right")
}



phyloseq_to_adonis <- function(physeq, distmat = NULL, dist = "bray", formula) {
  if(!is.null(distmat)){
    phydist <- distmat
  } else {
    phydist <- phyloseq::distance(physeq, dist)
  }
  metadata <- as(sample_data(physeq), "data.frame")
  
  # Adonis test
  f <- reformulate(formula, response = "phydist")
  adonis.test <- adonis(f, data = metadata)
  print(adonis.test)
  
  # Run homogeneity of dispersion test if there is only 1 variable
  if (length(formula) == 1) {
    Lake_Type <- metadata[,formula]
    beta <- betadisper(phydist, Lake_Type)
    disper.test = permutest(beta)
    print(disper.test)
  
    l <- list(
      dist = phydist, 
      formula = f, 
      adonis = adonis.test, 
      disper = disper.test
    )
    
  } else {
    
    l <- list(
      dist = phydist, 
      formula = f, 
      adonis = adonis.test
    )
  }
  return (l)
}
```

